### 2014-04-11 WSSSPE1_report/WSSSPE1_summary_report.tex
Author details for James H; tidied up double citation; fixed cite re environmental sustainability

\documentclass[11pt, oneside]{amsart}

\usepackage{amsmath}
\usepackage{amssymb}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}

\usepackage{color}
\usepackage{dcolumn}
\usepackage{float}
\usepackage{graphicx}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{subfigure}
\usepackage{psfrag}
\usepackage{tabularx}
\usepackage[hyphens]{url}
\usepackage{wrapfig}
\usepackage{enumitem}
\usepackage{multicol}
\usepackage{verbatim}

%\setcounter{secnumdepth}{3}
%\setcounter{tocdepth}{3}

\usepackage[bookmarks, bookmarksopen, bookmarksnumbered]{hyperref}
\usepackage[all]{hypcap}
% Don't use tt font for urls
\urlstyle{rm}

\definecolor{orange}{rgb}{1.0,0.3,0.0}
\definecolor{violet}{rgb}{0.75,0,1}
\definecolor{darkgreen}{rgb}{0,0.6,0}
\definecolor{cyan}{rgb}{0.2,0.7,0.7}
\definecolor{blueish}{rgb}{0.2,0.2,0.8}

\newcommand{\todo}[1]{{\color{blue}$\blacksquare$~\textsf{[TODO: #1]}}}
\newcommand{\katznote}[1]{ {\textcolor{magenta}    { ***Dan:      #1 }}}
\newcommand{\gabnote}[1]{ {\textcolor{cyan}    { ***Gabrielle:     #1 }}}
\newcommand{\nchnote}[1]{  {\textcolor{orange}      { ***Neil: #1 }}}
\newcommand{\manishnote}[1]{  {\textcolor{violet}     { ***Manish: #1 }}}
\newcommand{\davidnote}[1]{  {\textcolor{darkgreen}      { ***David: #1 }}}
\newcommand{\note}[1]{ {\textcolor{red}    { #1 }}}

% formatting commands
%   - consistent formatting of names of software tools and/or projects:
\newcommand{\toolname}[1] {\textsf{#1}}

% 15 characters / 2.5 cm => 100 characters / line
% Using 11 pt => 94 characters / line
\setlength{\paperwidth}{216 mm}
% 6 lines / 2.5 cm => 55 lines / page
% Using 11pt => 48 lines / pages
\setlength{\paperheight}{279 mm}
\usepackage[top=2.5cm, bottom=2.5cm, left=2.5cm, right=2.5cm]{geometry}
% You can use a baselinestretch of down to 0.9
\renewcommand{\baselinestretch}{0.96}

\sloppypar

\begin{document}

\title[]{Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1) 
\note{\scriptsize Authors, please add/check middle initials and affiliations}}

\author{Daniel~S.~Katz$^{(1)}$, Hilmar~Lapp$^{(2)}$, Ketan~Maheshwari$^{(3)}$, Frank~L\"{o}ffler$^{(4)}$, 
Matthew~Turk$^{(5)}$, Marcus~D.~Hanwell$^{(6)}$, Nancy~Wilkins-Diehr$^{(7)}$, James~Hetherington$^{(8)}$, 
James~Howison$^{(9)}$, Sou-Cheng~T.~Choi$^{(10)}$, Shel~Swenson$^{(11)}$,  Bruce~Berriman$^{(12)}$, Andy~R.~Terrel$^{(13)}$, Colin~Venters$^{(14)}$}

\thanks{{}$^{(1)}$ National Science Foundation, Arlington, VA, USA; Computation Institute, University of Chicago \& Argonne National Laboratory, Chicago, IL, USA; \url{d.katz@ieee.org}}
%
\thanks{{}$^{(2)}$ National Evolutionary Synthesis Center (NESCent),
  Durham, NC, USA; \url{hlapp@nescent.org}}
%
\thanks{{}$^{(3)}$ Argonne National Laboratory, Chicago, IL, USA; \url{ketan@anl.gov}}
%
\thanks{{}$^{(4)}$ Center for Computation and Technology, Louisiana State University, Baton Rouge, LA, USA; \url{knarf@cct.lsu.edu}}
%
\thanks{{}$^{(5)}$ Columbia Astrophysics Laboratory, Columbia University, New
York, NY, USA; \url{matthewturk@gmail.com}}
%
\thanks{{}$^{(7)}$ San Diego Supercomputer Center, University of California San Diego, La Jolla, CA, USA, \url{wilkinsn@sdsc.edu}}
%
\thanks{{}$^{(8)}$ Leader, Research Software Development Team, University College London; \url {j.hetherington@ucl.ac.uk}}
%
\thanks{{}$^{(9)}$ University of Texas at Austin; jhowison@ischool.utexas.edu}
%
\thanks{{}$^{(10)}$ NORC at the University of Chicago and the Illinois Institute of Technology, Chicago, IL, USA; \url{sctchoi@uchicago.edu}}
% 
\thanks{{}$^{(11)}$ xxx}
%
\thanks{{}$^{(12)}$ xxx}
%
\thanks{{}$^{(13)}$ xxx}
%
\thanks{{}$^{(14)}$ xxx}

\begin{abstract} %\note{Lead: Sou-Cheng T. Choi}
This document summarizes the First Workshop on Sustainable Software for Science:
Practice and Experiences (WSSSPE1). The content is based on the workshop
itself, as recorded in a set of collaborative notes taken during the workshop,
including discussion about the keynote presentations, three themed panels that
discussed papers accepted at the workshop, other discussion, cross-cutting
issues, use cases, and conclusions. It joins a nascent literature seeking to
understand what drives software work in science and how the reward systems of
science work, thereby shaping the type of software work undertaken, including the
extent to which developers are motivated to build software for the long-term,
for the use of others, and whether to work collaboratively or separately.
Unique perspectives were captured about issues such as comprehensive documentation,
software deployment, and development practices; software licenses; career paths
for developers; identifiers for software through a
mechanism such as a Digital Object Identifier or publication of a ``software
paper''; and evidence of software contribution and impact, recorded in online
systems such as source code repositories like GitHub. For truly sustainable
software, there should be no endpoint, as the software products continue to be
used and useful beyond a single institution, grant, and developer or
development team.

\end{abstract}

\maketitle

%\note{papers and presentations and URLs should be references - see
%  figshare-web in .bib file for URL style example}

\begin{comment}
People who have volunteered to work on this report: (with * for actual
contributors - \note{if you contribute, add a * by your name})

The workshop organizers:
\begin{itemize}
\item * Daniel S. Katz $\langle$\url{dkatz@nsf.gov}$\rangle$
\item Gabrielle Allen $\langle$\url{gdallen@illinois.edu}$\rangle$
\item Neil Chue Hong $\langle$\url{N.ChueHong@software.ac.uk}$\rangle$
\item Manish Parashar $\langle$\url{parashar@rutgers.edu}$\rangle$
\item David Proctor $\langle$\url{dproctor@nsf.gov}$\rangle$
\end{itemize}

Additional volunteers:
\begin{itemize}
% \item Chris Mattmann $\langle$\url{chris.a.mattmann@jpl.nasa.gov}$\rangle$
\item * Ketan Maheshwari $\langle$\url{ketancmaheshwari@gmail.com}$\rangle$
\item Marlon Pierce $\langle$\url{marpierc@iu.edu}$\rangle$
\item * Colin Venters $\langle$\url{C.Venters@hud.ac.uk}$\rangle$
\item Suresh Marru $\langle$\url{smarru@iu.edu}$\rangle$
% \item Lynn Zentner $\langle$\url{lzentner@purdue.edu}$\rangle$
\item Anne Elster $\langle$\url{anne.elster@gmail.com}$\rangle$
\item * Shel Swenson $\langle$\url{mdswenso@usc.edu}$\rangle$
\item * Andy Ray Terrel $\langle$\url{andy.terrel@gmail.com}$\rangle$
% \item Abani Patra $\langle$\url{abani.patra@gmail.com}$\rangle$
\item * Nancy Wilkins-Diehr $\langle$\url{wilkinsn@sdsc.edu}$\rangle$
% \item James Spencer $\langle$\url{j.spencer@imperial.ac.uk}$\rangle$
\item * Frank L\"{o}ffler $\langle$\url{knarf@cct.lsu.edu}$\rangle$
\item * James Hetherington $\langle$\url{j.hetherington@ucl.ac.uk}$\rangle$
\item * Sou-Cheng (Terrya) Choi $\langle$\url{sctchoi@uchicago.edu}$\rangle$
\item * James Howison $\langle$\url{james@howison.name}$\rangle$
\item * Bruce Berriman $\langle$\url{gbb@ipac.caltech.edu}$\rangle$
\item * Hilmar Lapp $\langle$\url{hlapp@nescent.org}$\rangle$
\item * Marcus D. Hanwell $\langle$\url{marcus.hanwell@kitware.com}$\rangle$
\item Lucas Nussbaum $\langle$\url{lucas.nussbaum@loria.fr}$\rangle$
\item * Matthew Turk $\langle$\url{matthewturk@gmail.com}$\rangle$
\end{itemize}

The original document is
\url{https://docs.google.com/document/d/1eVfioGNlihXG_1Y8BgdCI6tXZKrybZgz5XuQHjT1oKU/edit?pli=1#}
(but can no longer be edited).  Note that the original document has
comments in addition to text.
\end{comment}

%\pagebreak
%
%\section*{Executive Summary}
%
%\todo{1 page summary goes here}
%
%\pagebreak

\section{Introduction}

The First Workshop on Sustainable Software for Science: Practice and
Experiences (WSSSPE1,
\url{http://wssspe.researchcomputing.org.uk/WSSSPE1}) was held on
Sunday, 17 November 2013, in conjunction with the 2013 International
Conference for High Performance Computing, Networking, Storage and
Analysis (SC13, \url{http://sc13.supercomputing.org}).

Because progress in scientific research is dependent on the quality of and
accessibility to software at all levels, it is now critical to address many
challenges related to the development, deployment, and maintenance of reusable
software.
%\note{Ketan: What are the levels in scientific process where software
%is required. It might be interesting to outline these stages and types of
%software required in each stage.}
In addition, it is essential that scientists,
researchers, and students are able to learn and adopt software-related skills
and methodologies. Established researchers are already acquiring some of these
skills, and in particular a specialized class of software developers is
emerging in academic environments who are an integral and embedded part of
successful research teams. The WSSSPE workshop was intended to provide a forum
for discussion of the challenges, including both positions and experiences. The
short papers and debates have been archived to provide a basis for continued
discussion, and the workshop has fed into the collaborative writing of this
document. An estimate of 90 to 150 participants were present
throughout the day. Additional papers based on extended versions of workshop
submissions are expected. The level of interest in the workshop has led the
organizers, working with some of the submitters and attendees, to plan two
additional workshops (at the 2014 SciPy and SC14 conferences), to obtain
funds from the National Science Foundation and the Gordon and Betty Moore
Foundation to support these workshops, and to turn the
workshop website into a community website that can be used as a focus for
further discussion and progress. Additionally, a minisymposium that aims to
further explore some of the key issues raised in WSSSPE is co-organized by a
WSSSPE1 participant at the 2014 Society for Industrial and Applied Mathematics
(SIAM) Annual Meeting on ``Reliable
Computational Science.''

The goal of this paper is to summarize the contents of the workshop papers
and discussion, and identify issues that future WSSSPE events will address.
Before the WSSSPE1 workshop took place, the organizers self-published
a paper~\cite{WSSSPE1-pre-report} to document the process of
organizing and advertising the workshop, collecting and reviewing the
papers, and putting together the agenda. Section~\ref{sec:process} is
based on that publication. The remainder of this paper is based on the
workshop itself, as documented in a set of collaborative notes taken
during the workshop~\cite{WSSSPE1-google-notes}, including discussion
about the keynote presentations (\S\ref{sec:keynotes}), the three
panels (\S\ref{sec:devel}-\ref{sec:community}), 
cross-cutting issues (\S\ref{sec:cross-cutting}),
use cases (\S\ref{sec:use-cases}), and conclusions
(\S\ref{sec:conclusions}).  Additionally, the accepted workshop papers are listed in
Appendix~\ref{sec:papers}, and a partial list of workshop attendees is
in Appendix~\ref{sec:attendees}.

\section{Workshop Process and Agenda} \label{sec:process}

WSSSPE1 was organized in collaboration of the relatively small group
of five organizers and a larger peer-review committee with 36
members. This committee had early influence on the organization, e.g.,
on the specific call for papers.

The workshop call for papers included:

\begin{quote}
``In practice, scientific software activities are part of an ecosystem
where key roles are held by developers, users, and funders. All three
groups supply resources to the ecosystem, as well as requirements that
bound it. Roughly following the example of NSF's Vision and Strategy
for Software~\cite{NSF_software_vision},
the ecosystem may be viewed as having challenges related to:

\begin{itemize}[leftmargin=0.2in]
\item the development process that leads to new (versions of) software
\begin{itemize}[leftmargin=0.2in]
\item how fundamental research in computer science or
  science/engineering domains is turned into reusable software
\item software created as a by-product of research
\item impact of computer science research on the development of
    scientific software and vice versa
\end{itemize}
\item the support and maintenance of existing software, including
  software engineering
\begin{itemize}[leftmargin=0.2in]
\item governance, business, and sustainability models
\item the role of community software repositories, their operation and
  sustainability
\end{itemize}
\item the role of open source communities or industry
\item use of the software
\begin{itemize}[leftmargin=0.2in]
\item growing communities
\item reproducibility, transparency needs that may be unique to science
\end{itemize}
\item policy issues, such as
\begin{itemize}[leftmargin=0.2in]
\item measuring usage and impact
\item software credit, attribution, incentive, and reward
\item career paths for developers and institutional roles
\item issues related to multiple organizations and multiple countries,
  such as intellectual property, licensing, etc.
\item mechanisms and venues for publishing software, and the role of
  publishers
\end{itemize}
\item education and training''
\end{itemize}

\end{quote}

Based on the goal of encouraging a wide range of submissions from
those involved in software practice, ranging from initial thoughts and
partial studies to mature deployments, the organizers wanted to make
submission as easy as possible. The call for papers stated:

\begin{quote}
``We invite short (4-page) position/experience reports that will be used
to organize panel and discussion sessions. These papers will be
archived by a third-party service, and provided DOIs [Digital Object Identifiers].
We encourage
submitters to license their papers under a Creative Commons license
that encourages sharing and remixing, as we will combine ideas (with
attribution) into the outcomes of the workshop.  An interactive site
will be created to link these papers and the workshop discussion, with
options for later comments and contributions. Contributions will be
peer-reviewed for relevance and originality before the links are added
to the workshop site; contributions will also be used to determine
discussion topics and panelists. We will also plan one or more papers
to be collaboratively developed by the contributors, based on the
panels and discussions.''
\end{quote}

58 submissions were received, and almost all submitters used either
arXiv~\cite{arXiv-web} or Figshare~\cite{figshare-web} to self-publish
their papers.

A peer review process followed the submissions, where
the 58 papers received 181 reviews, an average of 3.12 reviews per
paper. Reviews were completed using a Google form, which allowed
reviewers to choose papers they wanted to review, then to provide
general comments and scores on relevance to the organizers and
to the authors. The review reports enabled the organizers to decide
which papers to associate with the workshop, and allowed the authors
to improve their papers.

The organizers decided to list 54 of the papers (see Appendix~\ref{sec:papers} for a list of these papers) as significant
contributions to the workshop, a very high acceptance rate but one
that is reasonable, given the goal of broad participation and the fact
that the reports were already self-published. The papers were
grouped into three main categories, namely \emph{Developing and
Supporting Software}, \emph{Policy}, and \emph{Communities}. Each
subject area was associated with a panel and discussion at the
workshop.

The workshop itself consisted of two keynote presentations and the
three panels/discussion sessions. The panels were organized based on a
classification of the workshop submissions into three categories,
following the themes of the call for papers as modified by the areas
of the submissions. Three to four representatives from each submission
category were appointed as panelists, and assigned to read a subset of
the papers in that category and then discuss them in the panel.

\section{Keynotes \note{lead: Sou-Cheng (Terrya) Choi}} \label{sec:keynotes}

The WSSSPE1 workshop began with two keynote presentations.

\subsection{A Recipe for Sustainable Software, Philip E. Bourne.} \label{sec:keynote1}

The first keynote~\cite{WSSSPE1-keynote1} was delivered by Philip
E. Bourne of University of California, San Diego.  Bourne is a basic
biomedical scientist but has also formed four software companies. He co-founded
PLOS Computational Biology~\cite{plos-web} and helped develop the RCSB
Protein Data Bank~\cite{pdb-web}.
He is working on automating three-dimensional visualizations of cell
contents and molecular structures, a problem that has not been solved
and when done, would serve as a key function of software in biomedical
sciences.

Bourne's keynote presentation was entitled ``A Recipe for Sustainable
Software,'' and developed based on his experiences.  He emphasized that
sustainability for software ``does not just mean more money from
Government'' (see also Section~\ref{sec:defining-sustainability}).  Other
factors to consider, he mentioned, include costs of production, ease
of maintenance, community involvement, and distribution channels.

In places, Bourne said, development in science has improved thanks to
open source and hosting services like GitHub~\cite{github-web}, but for the most part
remains arcane. He argued that we can learn much from the App Store
model about interfaces, ratings, and so on. He also mentioned
BioJava~\cite{biojava-web} and Open Science Data Cloud~\cite{osdc-web}
as distribution channels.  On a related note, Bourne observed a common
evolutionary pathway for computational biology projects, from data
archive to analytics platform to educational use, and suggested that
use of scientific software for outreach might be the final step.

Bourne shared with the audience a few real challenges he
encountered. First, also an anecdote, he has looked into
reproducibility in computational biology, but has concluded that ``I
have proved I cannot reproduce research from my own
lab''~\cite{Veretnik}.

Another problem he experienced was staff retention with respect to
private organizations which reward those combining research and
software expertise (the ``Google Bus''). However, he is a strong fan
of software sustainability through public-private partnerships. He
noted that making a successful business from scientific software alone
is very rare: founders overvalue while customers undervalue. He noted
that to last, an open source project needs a minimal funding
requirement even with a vibrant community --- goodwill only goes so
far if one is being paid to do something else.  He talked about grant
schemes of relevance in the U.S., particularly with regard to
technology transfer~\cite{sbir-web, fased-web}.

He also had problems with selling research software: the technology
transfer office in his university wanted huge intellectual property
reach through, whereby they would get a share of profits from drugs
developed by pharmaceutical companies who use the software.  He was
aware this was unrealistic but the technology transfer office insisted
for a while. He wants to push a one-click approach for customers to
purchase university-written software.

He then presented arguments on directly valuing software as
a research output alongside papers, a common discussion within this field.
He mentioned an interesting exploration of involving software
engineers in the review process of scientific code~\cite{peer-review-code}.

On the notion of \emph{digital enterprise}, where information
technology (IT) underpins the whole of organizational activities, he
contended that universities are way behind the curve. In
particular, he highlighted the separation of research, teaching, and
administration into silos without a common IT framework as a blocker
to many useful organizational innovations: ``University 2.0 is yet to
happen.'' He spoke of a circumstance where someone had used an
algorithm developed for computational biology in marketing.
The role of an institution is important in this space. He argued that
funders can help train institutions, not just individuals, in this regard.

He concluded with a reference to his paper~\cite{bourne_ten} and
argued that computational scientists ``have a responsibility to
convince their institutions, reviewers, and communities that software is
scholarship, frequently
more valuable than a research article,'' a point with which all authors
strongly agree.

\subsection{Scientific Software and the Open Collaborative Web, Arfon Smith} \label{sec:keynote2}

The second keynote~\cite{WSSSPE1-keynote2} was delivered by Arfon
Smith of GitHub.

Smith's talk started with an example from data reduction in Astronomy,
where he needed to remove interfering effects from the device. He
built a ``bad pixel mask,'' and realized that while it was persistent,
there was no way or practice of sharing this with the
data. Consequently many researchers repeated the same calculations. He
estimated that 13 person-years were wasted in this redundant
calculation.

``Why didn't we do better?''  Smith asked of this practice. He argued
this was because we were taught to focus on immediate research
outcomes and not on continuously improving and building on tools for
research. He then asked, when we do know better, why we do not act any
different. He argued that it was due to incentives and their lack:
only the immediate products of research, not the software, are valued.
He referenced Victoria Stodden's talk at
OKCon~\cite{okcon-stodden-talk} which he said argued these points
well.

C. Titus Brown~\cite{ged-web}, a WSSSPE1 contributor and participant,
argued that with regard to reusable software, ``we should just start
doing it.'' In this regard Smith replied that documentation should be
``treated as a first class entity.''  He noted that the open source
community has excellent cultures of code reuse, for example,
RubyGems~\cite{rubygems-web}, PyPI~\cite{pypi-web}, and
CPAN~\cite{cpan-web}, where there is effectively low-friction
collaboration through the use of repositories. This has not happened
in highly numerical, compiled language scientific software.  An
exception he cited as a good example of scientific projects using
GitHub is the EMCEE Markov Chain Monte Carlo project~\cite{emcee-web}
by Dan Foreman-Mackey and contributors.

He argued that GitHub's \emph{Pull Request} code review mechanism
facilitates such collaboration, by allowing one to code first, and
seek review and merge back into the trunk later.

``Open source is \ldots reproducible by necessity,'' Smith quoted
Fernando Perez~\cite{perez-open-src-reproducible}, explaining that
reproducibility is a prerequisite for remote collaboration.  He
pointed out that GitHub could propel the next stage of web
development, i.e., ``the collaborative web,'' following on from the
social web of Facebook.

In conclusion Smith reiterated the importance of establishing
incentive models for open contributions and tool builders, for
example, meaningful metrics and research grants such as~\cite{NSF_software_vision}. He urged computational scientists to
collaborate and share often their research reports, teaching
materials, code, as well as data by attaching proper licenses.

\section{Developing and Supporting Software \note{\scriptsize lead: Marcus Hanwell, contributor: Suresh Marru}} \label{sec:devel}

The panel on developing and supporting software examined the challenges
around scientific software development and support, mainly focused on research
groups that also produce code in various forms. There was widespread agreement
that developing and maintaining software is hard, but best practices can help.
Several participants added that documentation is not just for users, and writing
API documentation, tutorials for building and deploying software, along with
documented development practices can be very helpful in bringing  new developers
into a project efficiently.

\subsection{Research or Reuse?}

Software is developed for many different purposes, and the requirements can
vary significantly depending on the intended audience. Most end-users make use
of either a graphical user interface of some kind, or a command line that may
offer input and output formats for running the code and analyzing its output(s).
When discussing backward compatibility it is these various interfaces that are
discussed. For software that builds on other software frameworks it is the
API (application programming interface) that is most important, and this can
encompass issues such as the source and binary interfaces to the software
libraries developed---with each potentially having a high maintenance cost if
they are to remain compatible over many years. When using command line programs
it is generally the command line switches as well as the input/output formats
that could incur significant costs if they are changed.

There was discussion that backward compatibility is not always desirable, and it
can be very costly. This must be balanced with the aims of a given project, and how
many other projects depend on and use the code when backwards incompatible
changes are to be made. There are many examples in the wider open source
software world of strategies for dealing with this, and again best practices
can go a long way to mitigating issues around backwards compatibility. Many
projects live with sub-optimal code for a while, and may allow backwards
compatibility to be broken at agreed-upon development points, such as a major
release for a software library.

There were 13 articles about different experiences in this area, but little
about GUI testing, performance, scalability, or agile development practices.
There were several unique perspectives about issues such as managing API
changes, using the same best practices for software as data, and going beyond
simply ``slapping an OSI-approved license on code.''

\subsection{Communities and Career Paths}

Communities are extremely important in software projects, and both their
building and continued engagement need attention during the project life cycle.
Several of the submitted papers discussed how communities have been built around
projects, and what is needed to enable a project to grow. Among these are public
source code hosting, mailing lists, documentation, wikis, bug trackers, software
downloads, continuous integration, software quality dashboards, and of course,
a general web presence to tie all of these things together.

There was extended discussion about users not answering the questions of other
users. Several participants
offered counterexamples, such as mailing lists where developers do not participate
as much due to users being more active, or whether the ``core
team'' can end up setting unrealistic expectations by doing too much. Team Geek~\cite{opac-b1134063}
and Turk's paper on scaling code in the human dimension~\cite{Turk:2013:SCH:2484762.2484782}  discuss how development
lists that are welcoming to people actually have many more people contributing.

Recruiting and/or retaining personnel in this area is hard,
with one of the major reasons being no long-term career paths (especially when
compared with industry). This is further discussed in \S\ref{sec:career-tracks}.

%moved to cross-cutting issues
%
%How should software development be balanced with
%research? It is apparent that things are beginning to improve, such as
%incentives for software development in the form of altmetrics, tenure committee
%consideration, and NSF ``products'' vs ``publications.'' It was noted it can be very
%difficult to measure where people are using small bits of your code.

\subsection{Defining Sustainability for Scientific Software}

The question of what ``sustainable'' means in the context of software was
raised (see \S\ref{sec:defining-sustainability}.) The resources at \url{http://oss-watch.ac.uk/resources/ssmm} discuss what
to look for when choosing software for procurement, or reuse in further software
development. Regardless of the license and development model that will be used,
the future of the project must be considered. Even if a particular piece of
software satisfies today's requirements, will it continue to do so in the
future? Important questions include whether the supplier will still be around in five
years time, will it still care for its customers needs, will it be responsive to
bug reports and feature requests? Should you tie your investment to a single
supplier using a proprietary product, or ensure the project uses an OSI-approved
license, and can outlive any single entity if the software is still important.

What is the overarching goal of sustainability? Is it reproducible science,
persistence, quality, something else? How should success be measured in this
context? Is there some metric that can be used to determine when you have
achieved sustainable software, or is this an ongoing process with no clear
endpoint. For truly sustainable software there should be no endpoint, as the
software products continue to be used and useful beyond the single institution,
grant, and developer/development team. Sustainability must be addressed
throughout the project life cycle.

\subsection{Software Process, Code Review, Automation, Reproducibility}

The papers submitted to this panel, and several others, include lots of general
recommendations for processes, practices, tools, etc. One of the papers
suggested that a ``Software Sustainability Institute'' should be vested with
the authority to develop standardized quality processes, a central common
repository, central resources for services and consulting, a think tank of
sorts, and a software orphanage center. The idea of one common repository received
some resistance, with so many compelling alternatives available, e.g. Bitbucket or
GitHub. The point for centralization of communication/point of contact was seen as
reasonable, with the statement that ``vested with authority'' is perhaps too strong,
but ``providing tools if needed'' might be more appropriate.

What about actual software engineering principles, such as modularity and
extensibility? This is how industry maintains software, and ensures it continues
to be useful. Often, rewriting software is considered to be too costly, but
with a modular design it can be kept up to date. Extensibility is expected to keep
it relevant, if built into the project. One counterpoint raised by Jason Riedy
was that trying to take advantage of the latest and greatest hardware often
makes this painful, hence the lack of papers mentioning ``GPUs and exotic
hardware.''

The question of whether funders, such as the NSF, can mandate software plans in
much the same way as they do data management plans, was raised. Daniel Katz responded that
software is supposed to be described as part of the NSF data management plan,
and that in NSF's definition, data includes software. A comment from Twitter
(@biomickwatson) raised the issue that this requires reviewers and funders who
understand the answers that are given in these plans. Daniel Katz responded
that in programs focused on software or data this can be done effectively, but
agreed that in more general programs this is indeed a problem.

\subsection{Training Scientists to Develop and Support Software}

Some of the questions raised after the panel discussion are:

Rather than teaching developers about domain science, or domain scientists about
software development practices, why don't we teach both communities to
collaborate more effectively? Can this be done without teaching each side a
little of the other to enable communication, with a response that it really is
not two binary communities, but a spectrum with useful roles in the middle.
The question was raised if a developer can be effective without being part of the
domain community, with responses that this really depends on the specific
problem---translators and T-shaped people are important. Why aren't academic
communities taught how to evaluate cross-disciplinary work well?

What is the role for the growing field of team science? There
is overlap between the communities, with support for virtual organizations,
tool development, etc. How can we make time in an already crowded schedule to
introduce these topics to students? Should they be introduced through lab
classes as in ``real sciences''?

\subsection{Funding, Sustainability Beyond the First Grant/Institution}

Are there significant differences in projects that have been running for 1, 3,
5, or 10+ years? Are there shared experiences for projects of a similar stage
of maturity? It was noted that computing and communication have changed
significantly over the past decade, and many of the experiences are tied to
the history of computing and communication. See the history of GCC, Emacs, or
the Visualization Toolkit for examples. Others felt that computing has changed
less, but communication and the widespread availability of tools has. It was
noted that email lists, websites, chat rooms, version control, virtual and
physical meetings are all over 20 years old.

There was debate that while some of the basics of computing may be fairly
similar tools commonly used for computing have changed quite
significantly. Reference was made to Perl, which was commonly used, giving way
to whole new languages, such as Python, for gluing things together and how this
induces many students into entirely rewriting the scaffolding, leaving the old
to rot and the experiments to become non-reproducible as the tools change.
Jason Riedy stated that he had been guilty of this in the past. There was
discussion of this tendency along with the enormous differences in the speed and
ease of sharing---having to ship tapes around (GCC, Emacs, etc) as opposed to
the immediate sharing of the latest development online, using revision control
systems like CVS, Subversion, Git, Mercurial, Bazaar, etc.

The question was also posed as to whether the distinction between researcher
and developer is sensible, with James Hetherington commenting that in the UK
a more nuanced view of research software engineers
and researcher developers is examined. Should this be less of a contract relationship, and
more of a collaborative relationship? This is also at the core of the business
model that Kitware presented in its submission to the workshop. Are other
ingredients missing such as applied mathematicians? Should this be defined more
in terms of skill sets rather than roles and/or identities? This builds on the
comments from Vaidy Sunderam that scientists are generally good writers, and
have mathematical skills, so why can't they learn software engineering
principles?

Miller commented that all of the infrastructure that sits around a new
algorithm that we need to make it useful and sustainable requires different
skill sets than the algorithm developer. Friere commented that there are no
good career paths for people with broad skills, no incentives for them to
continue in these roles. There was debate around people doing what interests
them, and learning computing leaves people cold, but is it that it leaves the
people who find career paths in academia cold versus the full spectrum of
people involved in research? Is this also caused by poor teaching, or because the
benefits for doing this are perceived as too small? It could also be attributed to
their focus being on science, not software engineering, or do people with
the passion for software engineering in science simply have no viable career
path and either adapt or seek out alternate career paths?

\section{Policy \note{lead: Colin Venters, contributors: James Howison, Hilmar Lapp}} \label{sec:policy}

The Policy panel was designated to discuss the group of workshop
contributions dealing with the wide range of software sustainability
aspects that relate to establishing, promoting, and implementing
policies. Six papers presented frameworks for defining, modeling, and
evaluating software sustainability, the basis of establishing
policies. Four papers advocated mechanisms for more properly assessing
the impact of scientific software, and for crediting and recognizing
work that promotes software sustainability, all of which are
instrumental in effectively promoting policies that aim to change
current practices. Four papers discuss facets of implementing software
sustainability, and models of implementation across different facets.

\subsection{Modeling Sustainability}

The workshop submissions grouped under this section provide frameworks
for thinking about, researching and understanding which elements of sustainability
are important and how they relate. Although there is substantial
overlap between the frameworks, they have different emphases and
extents.  Each paper in this group included a definition of sustainability,
with many overlaps.  In essence the papers related the question to
change over time, an idea well captured in the definition used by the
UK's Scientific Software Sustainability Institute: ``software you use
today will be available---and continue to be improved and
supported---in the future,'' as quoted in~\cite{Venters_WSSSPE}. This
idea was also expressed as software that continues to serve its
users~\cite{Pierce_WSSSPE}.  For a more thorough discussion of definitions
of software sustainability see \S\ref{sec:defining-sustainability}.

One area in which there was not complete overlap was whether the word
(and thus the effort called for by the WSSSPE workshop) involved
environmental sustainability. Of course the word sustainability has
strong connotations from consideration of environmental issues,
evoking some mention of the areas in which software interacts with
overall environmental resource usage, particularly energy efficiency.
The two papers in this area which mentioned this~\cite{Venters_WSSSPE,Calero_WSSSPE} did so without
integrating that analysis into the question of software being around
long-term, suggesting that questions of environmental impact of
scientific software is a conceptually distinct area of inquiry.

One group of papers presented frameworks that were primarily about
characteristics of software artifacts, connecting with the long
discourse on software quality. This approach is realized in adjectives
that can be applied to pieces of software but might also extend to
describe software projects.  Thus Calero et al.~\cite{Calero_WSSSPE} propose adding elements to the ISO
standards for measuring software quality. Specifically, they propose an additional
dimension of quality they call ``perdurability'' or ability to ``last over
time'' made up of three characteristics: reliability, maintainability
and adaptability. This shows some overlap with the framework employed
by Venters et al.~\cite{Venters_WSSSPE} who employ the adjectives of
``extensibility, interoperability, maintainability, portability,
reusability and scalability,'' anticipating the sorts of work that
people would need to do with a software artifact in the future, ``as
stakeholders requirements, technology and environments evolve and
change.'' Venters et al. argue that these choices are really made very
early because they relate to the architecture of the software and
involve trade offs that ought to be analyzed alongside other
attributes of the software artifact. Lenhardt et
al~\cite{Lenhardt_WSSSPE} compare the software lifecycle to the data
lifecycle to argue for the inclusion of metadata throughout a piece of
software's life (discussing, for example, how it has been built and
tested and what ``data in'' and ``data out'' has been considered). In
addition, their analogy suggests that the software lifecycle might add
a phase of ``preservation'' and draw on understandings of what that
involves from studies of data.  In sum, then, these frameworks focused
on what needs to be accomplished to have more sustainable software.

A second theme in these papers was the continued availability of
resources to accomplish the goals of sustainability. The elements of
these frameworks focused far more on the organization of a software
project than they did on characteristics of the artifact itself
(although it is certainly true that the adjectives discussed above
could be applied to a software project). For example Pierce et
al.~\cite{Pierce_WSSSPE} focus on the way that the project is run,
particularly in terms of how those involved communicate and jointly
set priorities, a process they call governance. In particular they
argue that because sustainability relates to having ongoing resources,
governance must be open to receive diverse input (by occurring online,
asynchronously, and publicly) and thus have the potential to
``transform passive users into active contributors''. They argue that
the Apache Software Foundation's incubation process teaches this and
could be learned from by projects throughout scientific software.
Katz and Proctor~\cite{Katz_WSSSPE} also discuss governance,
describing two modes: ``top-down'' and ``bottom-up'' governance.  They
place governance alongside questions of the artifact (``technical''),
and questions of the origins of resources, in terms of who is funding
the work surrounding a piece of software (``the political'') and the
manner in which resources come to the project initially (commercial,
open source, closed partnerships, grant funded) and long-term (all
four plus paid support).

Frameworks also concerned themselves with the context in which software
projects exist, moving in abstraction from the software artifact
itself to the organization of its production and
the shape of the space in which an artifact or project exists.
These frameworks take the form of contingency theories in that
they outline a different set of challenges and argue that different
project organizations (and presumably artifact attributes) are
necessary to persist longterm in spaces with particular
characteristics.  Katz and Proctor~\cite{Katz_WSSSPE} propose thinking
of this idea of space in terms of three axes: temporal (long or short
term needs), spatial (local or global use) and purpose (specific to
general).  They propose that different project organizations will be
needed in different locations and argue that we should concentrate
research to understand those connections. Weber et
al.~\cite{Weber_WSSSPE} describe their spaces by analogy with natural
ecosystems as ``niches'' which sustain particular pieces of
software. They define ``a software niche as the set of technical
requirements, organizational conditions and cultural mores that
support its maintenance and use over time.'' They call for better
understanding and modeling of niches (as well as further exploration
of the usefulness of ecosystem metaphors).

\subsection{Credit, Citation, and Impact}

How work on scientific software is recognized and rewarded strongly influences the motivation for particular kinds of work
on scientific software. A recurring theme of the panel discussion was
that software work in science is
inadequately visible within the reputation
system underlying science; in other words it often doesn't ``count''. In his paper for this workshop, Katz placed
software work along with other ``activities that facilitate science
but are not currently rewarded or
recognized''~\cite{Katz_WSSSPE}. Priem and Piwowar argued for the need
to ``support all researchers in presenting meaningful impact evidence
in tenure, promotion, and funding applications.''~\cite{Priem_WSSSPE}.
Kneply et al. argued that the visibility of software that supported a
piece of science ``can have detrimental effects on funding, future
library development, and even scientific
careers.''~\cite{Knepley_WSSSPE}.

These papers, and the discussion at the workshop, join a nascent
literature seeking to understand what drives software work in science
and how the reward systems of science thereby shape the type of
software work undertaken, including the extent to which developers are
motivated to build software for the long-term, for the use of others
and whether to work collaboratively or
separately~\cite{howison_incentives_2013, howison_scientific_2011,
  bietz_synergizing_2010}. Software work is not only motivated by
direct citations, but the visibility of software work in the
literature is important to those who write software used in science.

Papers and discussion concentrated on three overarching questions: How ought software
work to be visible, what are the barriers to its visibility, and what
can be done to make it more visible?

Most of the papers in this area focused on visibility of software in
scientific papers, since scientific papers are the most widely
accepted documentation of achievement in science. It was noted that
there appear to be no widely accepted standards on how the use of
software towards a paper ought to be mentioned, and that journals,
citation style guides and other guides to scientific conduct are vague
about how to describe software. To address this, papers advocated the
need for a fixed identifier for software, either directly through a
mechanism such as a Digital Object
Identifier~\cite{Katz_WSSSPE,Knepley_WSSSPE} or via a published paper
written to document the software (and perhaps its creation), a
``software paper''~\cite{Chue_Hong_WSSSPE}. However, as was pointed
out during the panel discussion, one of the problems with papers as
the cited product is that their author list is fixed in time, which
discourages potential contributors who are not on the original author
list from designing incremental improvements as integration work
rather than separate (and hence possibly rewritten) software
products~\cite{howison_incentives_2013}.

Another approach is to reduce the difficulty of citing all software
underlying a research paper. For example, scientists often work with
software that itself wraps other software, leading to attribution
stacking that can make it non-obvious or even difficult to determine
what attributions would be appropriate. Knepley et al.~\cite{Knepley_WSSSPE} approach this by
proposing a mechanism by which the software itself, after it has run,
provides the user with a set of citations, according to the pieces of
code actually executed. They describe a prototype implementation whereby
the citations are embedded in libraries and reported along with the
results, via a command-line interface~\cite{Knepley_WSSSPE}. Discussion
highlighted the difficulty that attempting to acknowledge the
contributions of all pieces of dependent code within a paper faces the
difficulty of creating very long citation lists, straining the analogy
of code used to papers cited. Katz approaches this issue by proposing
a system of ``transitive credit,'' recording dependencies and relative
contributions outside particular papers, relieving authors from the
responsibility of acknowledging each and every dependency. Instead
authors would acknowledge the percentage contribution of the software
they used directly and an external system would then be able to
recursively allocate that credit to those who had provided
dependencies. Finally Priem and Piwowar argued that machine learning
techniques could examine the body of published literature and extract
mentions of software, coping with the multitude of informal ways in
which authors mention software they used~\cite{Priem_WSSSPE}.
A point raised in the panel discussion was that instead of asking users to improve
their software citation practices, one can also ask how software
projects can better monitor the literature to improve their
ability to show impact. For example, the NanoHub project scans the literature using keywords and names
of known users to discover papers that are likely to have used their
software and platform, and assigns graduate students to read each paper,
highlighting mentions of software use and sometimes following up with the authors to
identify stories for demonstrating impact.  A process for tracking
software-using publications with the goal of increasing impact visibility is now described at \url{http://publications.wikia.com}.

Potential visibility, and thus acknowledgement of scientific software
products is not restricted to publications. Another key location for
visibility is in the grant funding
process, and as emphasized by NSF representatives at the meeting,
recent changes to grant proposal and reporting formats now allow both
applicants and awardees to report and highlight software products as
much as publications. Nonetheless, whether peer review
panels would value these contributions in the same way as
publications remains to be seen.

Priem and Piwowar argued that assessing the impact of software work
requires looking beyond publications, including evidence of contribution and impact recorded in social
coding-oriented resources such as Github, and conversations about software
in issue trackers, mailing lists, twitter and
beyond~\cite{Priem_WSSSPE}. In keeping with a principle of
the ``altmetrics'' approach, they advocate that scholars should have
resources that empower them to tell their own stories in the
manner most appropriate for them and their audiences.

\subsection{Implementing Policy}

The workshop contributions in this group were concerned with the
aspect of how implementation of best practices and other
recommendations for improving scientific software sustainability could
be promoted. Specifically, if scientific software is to become more
sustainable, corresponding policies and guidelines need to be such
that the scientific community can follow and implement them. This is
considerably more challenging that it might seem at first, because in
the reality of science today resources, both financial and personnel,
that could be devoted to implementation are very limited, and the
reward system does not encourage scientists to do so. Furthermore,
implementing sustainability-targeting policies and guidelines often
takes a variety of specialized software engineering expertises, which
are not necessarily found in a single engineer, and much less so in a
domain scientist cross-trained in programming. Adding to the policy
implementation challenges, applicable sustainability-promoting
practices and guidelines will change through a software project's
lifecycle, in particular as it gains maturity.

Two of the papers in this group focus on specific facets of software
design that are important factors in a project's sustainability but
are often addressed only late in the scientific software development
cycle, if at all: Krintz et al.~\cite{Krintz_WSSSPE} look at API
governance, and Heiland et al.~\cite{Heiland_WSSSPE} discuss maturity
models for software security. The other two papers discuss
implementation strategies for science from the perspective of
facilitating many or all facets of sustainanability-oriented software
design: Blanton and Lenhardt~\cite{Blanton_WSSSPE} contrast
large projects that have software infrastructure development built-in,
with cross-training domain scientist PIs in software engineering best
practices. Huang and Lapp~\cite{Huang_WSSSPE} discuss how various
specialized software engineering skills could be turned into shared
instrumentation with low barriers to access.

Krintz et al.~\cite{Krintz_WSSSPE} describe how in an era in which
computing frequently takes place in a distributed cloud, the control
over digital resources is increasingly shifting from physical
infrastructure to APIs, in particular web-service APIs. Yet, as Krintz
et al. observe, unlike for physical IT infrastructure in data centers,
science communities have developed very little in the way of practices and
technology for API governance, referred to by Krintz et al. as the
``combined policy, implementation, and deployment control''. Web APIs
can and do change, sometimes quite frequently. Raising the need to
port dependent applications. The effort required for porting is
notoriously difficult to estimate, making it nearly impossible for IT
organizations to assess and thus properly manage the impact of API
changes. To address this, Krintz et al. propose a mechanism that
evaluates the porting effort between two versions of a web-service API
in a formal and automated way. To analyze the porting effort, they
divide API changes into \emph{syntactic similarity}, the changes in
inputs and outputs, and into \emph{semantic similarity}, the changes
in the API's behavior. In initial tests, their method showed good
congruence with human developers in scoring porting effort, offering
the possibility that API governance can become as solid a part of
scientific IT management as data center infrastructure management is
today.

Many facets of engineering for software sustainability strongly depend
on the maturity level of the software. However, the maturity level of
a software project meant to be sustained changes tremendously over its
development life cycle, and the eventual maturity level is often
difficult to predict during initial development. Using software
security as their case study, Heiland et al.~\cite{Heiland_WSSSPE}
discuss how Maturity Models can be used to formalize best practices
appropriate for the different levels of maturity that a software
project may go through over its lifecycle. Cybersecurity is also an
example of a sustainability-relevant aspect in software design that is
rarely given due diligence in science. In particular in industry,
cybersecurity best practices for different stages of life cycle and
maturity have been formalized as Software Security Maturity Models,
and are widely used, yet awareness of these among scientific software
development communities remains low. In providing a path to tightening
security practices as software matures, such models align with the
objective of providing implementation approaches that the scientific
community can actually follow.

API governance and cybersecurity measures appropriate for a project's
maturity are all but two facets of sustainability-oriented software
development. Others include user-centered interface design, test
engineering, dependency management, and deployment operations. Each of
these facets require specialized skills and training in software
engineering. How can the implementation of best practices and
guidelines along many or all of these different facets be facilitated
in scientific software development? Blanton and
Lenhardt~\cite{Blanton_WSSSPE} contrast two models. In one, the time
and personnel devoted to software engineering is ``co-funded'' with
the driving science project. This typically implies large multi-year
collaborative projects that to succeed require significant software
infrastructure to be built, and which thus have the funding to support
one or several software engineers. The sustainability of such projects
then depends on sustaining the funding. In the other model, domain
scientists also take the role of software developers, whether by
necessity such as funding limitations, or due to cross-disciplinary
professional interests. For this to result in sustainable software
products, the domain scientists need to be (or become) cross-trained
in software engineering standards and best practices.

In practice, there are example for both extremes of the spectrum. For
example, in the life sciences the iPlant Collaborative~\cite{iPlant},
the Galaxy Project~\cite{Galaxy}, and \toolname{Qiime}~\cite{Qiime} are large
multi-year projects with significant software infrastructure funding
and deliverables. Typically though these are the exception rather than
the rule, and particularly in the long tail of science the
scientist-developer predominates, even for software that is widely
used or crosses domain boundaries such as
\toolname{rOpenSci}~\cite{rOpenSci}. For better training domain scientists at
least in basic software engineering best practices, initiatives such
as Software Carpentry~\cite{SoftwareCarpentry} have demonstrated how
this could be achieved at scale.

However, there may also be a middle ground between the two
extremes. Huang and Lapp~\cite{Huang_WSSSPE} propose a Center of
Excellence model that leverages economies of scale to make software
engineering experts and their skills accessible to the long tail of
science. As Huang and Lapp discuss, this model could effectively turn
the utilization of software engineering expertise from a complex human
resource recruitment and management challenge, to buying time, when
and to the extent needed, on shared instrumentation. There is
precedence to using such a model to lower access barriers for long
tail science, in particular for new experimental technologies.  For
example, although the acquisition and operation of next-generation
high-throughput DNA sequencers requires substantial investments of
money, time, and expertise, the establishment of ``core facilities''
on many university campuses has made these technologies accessible to
a wide swath of scientists, with transformative results for science.

One of the important conclusions from this group of papers is that
creating sustainable software requires paying attention not to one or
two, but to several different facets of software engineering, each
with corresponding best practices and standards of excellence. Even if
a science project requires and has funding for a full software
engineering FTE, what the project really needs could be fractions of
different engineers with different specialty training. The vast
majority of long tail science projects lacks the funding for even one
full software engineer, let alone one who combines expertise in all of
the applicable facets of engineering. Some scientists in the long tail
will have the professional interests to cross-train enough in software
engineering to be successful with creating sustainable software, but
it is unrealistic to expect this expertise and interest of all or even
the majority. This is where a software engineering center of
excellence could provide a critical resource by enabling scientists in
the long tail to complement their resources and expertise with facets
that are missing, but which, if applied at the right time, would
improve the chances of a software product to become sustainable. Such
complementary expertise also need not be restricted to software
engineering in the strict sense; for example, it could consist of
community building, leadership, and support for some period of
transition to sustainability.

In summary, implementing software sustainability practices on a
broader basis requires on the one hand the development of guidelines
and practices that are suitable for the exploratory research context
in which most scientific software is created, and on the other hand a
skilled workforce trained in a variety of software engineering facets
and community building. The capabilities afforded by such a workforce
need to be accessible not only to large projects with sufficient
funding to provide competitive employment, but also to the many
smaller projects in the long tail of science. Sufficiently
cross-training domain scientists could be one way to achieve this;
another, complementary approach is to instrument the necessary
capabilities so they can be shared.

\section{Communities \note{\vspace{-1ex}\tiny lead: Matthew Turk, contributors: Nancy Wilkins-Diehr, Chris Mattmann, Suresh Marru, Frank L\"{o}ffler, Andy Terrel}} \label{sec:community}

%There were a number of papers categorized under the Communities banner and so
%two presenters each summarized half of the submissions in this area.
Across all talks and papers submitted, authors implicitly and explicitly
recognized the concept of ``communities'' as a driving and unifying force
within software projects.  Despite that, the actual nature of the communities,
the incentive structures and bind them together, the infrastructure they
utilize for communication and even the types of individuals that make up those
communities were different.  Some of these communities were primarily composed
of members of industry, some were funded and driven by individuals focused on
developing software as opposed to utilizing it, and others were primarily
composed of scientist practitioners, and were true communities of practice.

These varying structures and compositions result in differing modes of
interaction within communities, styles of development, and the structure of
planning for future development of software.  In this section, we summarize
the different types of communities in scientific software as well as the
resultant impact on sustainability and development of functionality.

\subsection{Communities}

%This collection of papers was summarized by Karen Cranston.

Drawing on experiences from high-energy physics, Vay et al.~\cite{Vay_WSSSPE} proposed
developing teams of technical specialists able to overcome a lack of
coordination between projects.  Maximizing scientific output requires
maximizing the usability of the scientific codes while minimizing the cost of
developing and supporting those codes.  This included targeting different
architectures for their software to be deployed, as well as coordination
between technically-focused individuals and usage of a common scripting
language between projects.  Instead of fragmenting the development of
simulation codes across institutions, the paper suggests that a cohesive
strategy reducing duplication and increasing coordination will broadly increase
the efficiency across institutions.  The approach proposed is of de-fragmenting
the existing ecosystem in a non-disruptive way.

Maheshwari et al.~\cite{Maheshwari_WSSSPE} focuses on ``technology
catalysts'' and their role in the modern scientific research process. A technology
catalyst is an individual with knowledge of
technological advancements, tasked with user engagement to create
scientific or engineering applications, using suitable tools and techniques
to take advantage of current technological capabilities.
%
One of the tasks of technology catalysts is to seek community collaborations for new
applications and engage users, thus benefiting both science, by effective
running of scientific codes on computational infrastructure, and technology,
by conducting research and seeking findings for technology improvement.
The particular engagements described in the paper came up from the lead author's work as
a postdoctoral researcher at Cornell and Argonne, where interaction with the scientific
communities in both institutions resulted in these collaborations.

At NESCent, a combination of in-house informatics individuals and domain
scientists collaborate to develop software to study evolutionary science.  The
report~\cite{Cranston_WSSSPE} studied the success of a ``hackathon'' model
for development, where short-form, hands-on events combining users,
researcher-developers and software engineers targeted specific code
improvements. From this experiment, the authors identified several key
outcomes as well as lessons-learned: specifically, the co-localization of
developers was seen as having a strong impact, enabling casual conversation
that led to discrete outcomes.  The formation of the discussion mailing list,
specifically in response to the social capital built at the hackathon, was seen
as spurring on longer-term benefits to the community and fostering
sustainability.

Hart et al.~\cite{Hart_WSSSPE} addresses the success of the rOpenSci project in developing
collaborations supporting tool development for Open Science.  This software
collective, organized around the statistical programming environment R,
develops access mechanisms for data repositories and attempts to reduce the
barrier to entry for individuals wanting to access data repositories and study
the data contained therein.  The collective fosters direct collaboration
between individuals and data providers, designed to ``train academics in
reproducible science workflows focused around R.''  Two central challenges
are engagement of existing users within ecology and
evolutionary biology (EEB), and how the community can make inroads and traction
in other disciplines. Currently, the collective is exploring addressing these
challenges through use of social media, holding workshops, and hackathons.
This helps to both raise the profile of the collective within EEB and in other
domains.  However, the overarching challenge identified in the paper was that
of incentivizing maintenance of software, which is difficult in academia.

%\subsection{Communities Part 2}

%Nancy Wilkins-Diehr summarized 6 papers in this part.

Christopherson et al.~\cite{Christopherson_WSSSPE} outlines the degree to which research relies on high
quality software. There are often barriers and a lack of suitable incentives
for researchers to embrace software engineering principles. The Water Science
Software Institute is working to lower some of these barriers through an Open
Community Engagement Process. This is a 4-step iterative development process
that incorporates Agile development principles.

\begin{itemize}
\setlength{\itemindent}{0.5in}
\item[Step 1:] Design - thorough discussion of research questions
\item[Step 2:] Develop working code
\item[Step 3:] Refine based on new requirements
\item[Step 4:] Publish open source
\end{itemize}

Christopherson reports on the application of Steps 1-3 to a
computational modeling framework developed in the 1990s. Step 1 was
a 2-day, in-person specifications meeting and code walkthrough. Step 2 was a
5-day hackathon to develop working code, and Step 3 was a 3-day hackathon to refine
the code based on new requirements. The team worked on small, low-risk units of
code. It was challenging, revealed unanticipated obstacles, programmers had to
work together, and experimentation was encouraged.
%
The paper recommended: start small and gradually building toward more complex
objectives, consistent with Agile development; develop consensus before
coding, by repeating step 1 before all hackathons; ensure newcomers
receive orientation prior to the hackathon, such as a code
walkthrough or system documentation; and co-locate collaborators
whenever feasible.

%Refactor before adding
%new functionality. Approach development as a learning experience. Welcome
%experimentation, and treat mistakes as a natural part of the learning process.
%Repeat Step 1 activities before all hackathons to develop consensus before
%coding. This allows coding to be the focus of subsequent hackathons. In higher
%risk situations, provide additional time for Step 1 activities. The
%Christopherson team recommends a minimum of two months. 
%Ensure any newcomers
%receive some form of orientation prior to the hackathon, such as a code
%walkthrough or system documentation. Co-locate rather than collaborating
%remotely whenever feasible.

Pierce et al.~\cite{Pierce2_WSSSPE} described how science gateways can provide a user-friendly
entry to complex cyberinfrastructure. For example, more than 7,000
biologists have run phylogenetic codes on supercomputers using the \toolname{CIPRES}
Science Gateway in 3 1/2 years. Over 120 scientists from 50 institutions used the
\toolname{UltraScan Science Gateway} in one year, increasing the sophistication of analytical
ultracentrifugation experiments worldwide. The new Neuroscience Gateway (\toolname{NSG})
registered 100+ users who used 250,000 CPU hours in only a few months.

Gateways, however, need to keep operational costs low and can often make use of
common components, such as authentication, application installation and
reliable execution and help desk support. Science Gateway Platform as a Service (\toolname{SciGaP}) delivers
middleware as a hosted, scalable third-party service while domain developers
focus on user interfaces and domain-specific data in the gateway. 
While \toolname{SciGaP} is based on the \toolname{Apache Airavata} project
and the \toolname{CIPRES} Workbench Framework, community contributions are encouraged
by its open source, open governance and open operations policies. The
goal is robust, sustainable infrastructure with a cycle of development that
improves reliability and prioritizes stakeholder requirements. The project is
leveraging Internet2's \toolname{Net+} mechanisms for converting \toolname{SciGaP} and its gateways
into commodity services.

Zentner et al.~\cite{Zentner_WSSSPE} describes experiences and challenges with the large
\toolname{nanoHUB.org} community, where community is defined as a ``body of persons of
common and especially professional interests scattered through a larger
society.'' Support is challenging because of the diversity of
viewpoints and needs. The group constantly examines its policies to determine
whether they are indirectly alienating part of the community or encouraging
particular types of use.
%
\toolname{nanoHUB}'s 10-year history with over 260,000 users annually provides a lot of
data to analyze. 4000 resources contributed by 1000 authors. \toolname{nanoHUB} serves
both the research and education community and the contribution model allows
researchers to get their codes out into the community and in use in education
very rapidly. The primary software challenges are twofold - support for the
\toolname{HUBzero} framework and challenges related to the software contributed by the
community.

The group has learned that community contributions are maximized with a
tolerant licensing approach. \toolname{HUBzero} uses an LGPLv3 license so contributors can
create unique components and license as they choose. If they make changes to
source code, the original license must be maintained for redistribution. As far
as contributed resources, these must be open access, but not necessarily open
source. This allows contributors to meet the requirements of their institutions
and funding agencies. Quality is maintained via user ratings. Documentation is
encouraged and nanoHUB supplies regression test capabilities, but the user
community provides ratings, poses questions and contributes to wishlists and
citation counts, all of which incentivize code authors.

Terrel~\cite{Terrel_WSSSPE} describes support for the Python scientific community through
two major efforts: the SciPy conference and the NumFOCUS foundation. 
Since software sustainability relies on contributions from all sectors of the user
community, these efforts support these sectors, and help develop and mature Python.
%
The reliance
on software in science has driven a huge demand for development, but this
development is typically done as a side effort and often in a rush to publish
without documentation and testing. While the software is often created by academics, software support
can fall to industrial institutions. SciPy brings together industry,
government, and academics to share their code and experience in an open
environment.
%
NumFOCUS in a non-profit that promotes open, usable scientific software while sustaining the
community through educational programs; collaborative
research tools and documentation; and promotion of high-level languages,
reproducible scientific research, and open-code development. Governance
is a loose grantor-grantee relationship with projects, allowing funds
to be placed in the groups' accounts. This has raised money to hire
developers for open code development, maintain testing machines, organize the
PyData conference series, and sponsor community members to attend conferences.

L\"{o}ffler et al.~\cite{Loffler_WSSSPE} describes the Cactus project,
 which was started in 1996 by
participants in USA Binary Black Hole Alliance Challenge. Cactus has a flesh (core)
and thorns (modules) model, a community-oriented framework that allows
researchers to easily work together with reusable and extendable software
elements. Modules are compiled into an executable and can remain dormant,
becoming active only when parameters and simulation data dictate. Users can
use modules written by others or can write their own modules without changing
other code. The community has grown and diversified beyond the original science
areas.

The paper points out 4 keys to sustaining the community: modular design,
growing a collaborative community, career paths and credit. In modular design,
the Cactus project went far beyond standard practices of APIs. Domain specific
languages (DSLs) allow decoupling of components - for example I/O, mesh
refinement, PAPI counters, and boundary conditions abstracted from science
code. In academia, publications are the main currency of credit. Because the
project connects code developments to science, the work is publishable and
modules are citable. Because of the open source, modular approach, programmers
can see the impact of their contributions and often continue work after
graduation. Career paths remain a challenge, however. Tasks that are essential
from a software engineering perspective are often not rewarded in academia. The
best programmers in a science environment often have multidisciplinary
expertise. This also is not rewarded in academia.

Wilkins-Diehr et al.~\cite{Wilkins-Diehr_WSSSPE} describes an NSF software institute effort to build a
community of those creating science gateways for science. These
gateways, as described in some of the other papers in this section, can be
quite capable and can have strong scientific impact.
Challenges are similar to those highlighted by other papers in this section:
the conflict between funding for research vs infrastructure and the
challenges around getting academic credit for infrastructure. 
%
The authors observe that  development is
often done in an isolated hobbyist environment. Developers are unable to take
advantage of similar work done by others, isolation even when projects have
common needs. But often projects struggle for sustainable funding because they
provide infrastructure to conduct research and many times only the research is
funded. Gateways also may start as a small group research project, taking off in
popularity once they go live, without any long term plans for sustainability.
Subsequent disruptions in
service can limit effectiveness and test the limits of the research community's
trust. %The impact of gateways can be increased substantially if we understand
%what makes them successful.

Recommendations from an early study of successful gateways include:
1) Leadership and management teams should design governance to
represent multiple strengths and perspectives, plan for change and turnover in
the future, recruit a development team that understands both the technical and
domain-related issues, consider sustainability and measure success early and
often. 2) Projects should hire a team of
professionals, demonstrate credibility through stability and clarity of
purpose, leverage the work of others, and plan for flexibility. 3) Projects should
identify an existing community and understand the communities' needs
before beginning, then adapt as the communities' needs evolve. 4) Funders,
should consider the technology project lifecycle, and design solicitations 
to reward effective planning, recognize the benefits and
limitations of both technology innovation and reuse, expect adjustments during
the production process, copy effective models from other industries and
sectors, and encourage partnerships that support gateway sustainability.

%Through a business incubator type approach, the institute plans to provide a
%variety of services that could be shared amongst projects. Consultation and
%resources on topics such as business plan development, software engineering
%practices, software licensing options, usability, security and project
%management as well as a software repository and hosting service will be
%available. Experts will be available for multi-month assignments to help
%research teams build their own gateways, teaching them how to maintain and add
%to the work after the collaboration ends. Forums, symposia and an annual
%conference will connect members of the development community. A modular,
%layered framework that supports community contributions and allows developers
%to choose components will be delivered and finally workforce development
%activities will help train the next generation for careers in this cross-
%disciplinary area and build pools of institutional expertise that many projects
%can leverage. Shared services and forums can dramatically reduce the cost of
%building and sustaining gateways. Workforce development can encourage
%technology professionals to remain in the sciences.

\subsubsection{What are communities?}

The workshop did not directly answer the question ``What are communities?'' but
instead a number of different answers were indirectly presented, through the
depiction of individuals and stakeholders in different aspects of the
scientific software lifecycle.  In broad strokes, however, scientific software
communities were generally accepted as consisting of individuals, often but not
always composed of scientist practitioners, that were working with some degree
of coordination toward a common goal enabled by software.

The discussion of development-focused communities centered around describing
methods of interaction between individuals and the scientific software.  The
first type of interaction was the development of a specific piece of software,
the second was a particular domain or discipline, and the final primary
type of interaction was around the development of applications built on a
particular piece of software that was perhaps developed by another group.  As
an example, in \cite{Hanwell_WSSSPE}, the community described is comprised of
both the for-profit company \toolname{Kitware} and the users and contributors to their
software packages such as \toolname{VTK}.  This structure, of the centralized development
of core infrastructure around which communities of individuals applying that
infrastructure and developing applications utilizing it, was similarly
reflected in \cite{Terrel_WSSSPE}, where the core scientific python ecosystem
is supported by a non-profit entity that fosters community investment in that
ecosystem.  In many ways, these two organizations (\toolname{Kitware} and \toolname{NumFOCUS})
attempt to cross domain boundaries and provide support for both the
infrastructure and application sides of community building.

\subsubsection{Measuring community}
How might a project know when it has built a sustainable community?  How might an outsider be able to assess the activity and sustainability of that community?  These questions have been partially addressed in the literature.  For example, a key metric in online communities in general is the cross-over point where there are more external contributors than internal ones. Richard Millington's book ``Buzzing Communities'' does an excellent job of outlining these measures, drawing on online communities research in an accessible manner~\cite{millington_buzzing_2012}. Some participants in the workshop have since prepared online materials outlining current and future practices in measurement of scientific software and its ecosystem\footnote{\url{https://sites.google.com/site/scientificsoftwaremetrics/}}. \katznote{please change to cite}

\subsubsection{Additional Resources for learning about software communities}

Scientific software communities were viewed as a subset of software communities
as whole.  As such, resources applicable to generic software communities --
such as open source and proprietary technology companies -- can be used as
input and as guiding understanding of how to steward and develop scientific
software communities.  Because incentive structures are different in industry
and volunteer-based open source communities, these can provide rough guidelines
but not necessarily identically applicable.  The analogy between corporations
and scientific investigators (particularly in terms of competition, cooperation
and competitive advantage) has been explored in the literature below, but
because of the different incentive structure the analogy is not universally
true.

The literature below, suggested by attendees, addresses both non-scientific
software projects, as well as scientific projects.  The selections address both
descriptive and prescriptive approaches to communities. 
\begin{itemize}
\item \cite{howison_scientific_2011, howison_incentives_2013}
\item \cite{citeulike:11831265}
\item \cite{citeulike:7888211}
\item \cite{citeulike:478633}
\item \cite{opac-b1134063}
\item \cite{Trapani:2011}
\item \cite{Allsopp:2012}
\item \cite{1749-4699-6-1-015010}
\item \cite{Turk:2013:SCH:2484762.2484782}
\end{itemize}

\subsection{Industry \& Economic Models}

Several papers presented discussed the connection between industry and
scientific software, from the perspective of both integrating efforts between
the two and sustaining long-term development.

Hanwell et al.~\cite{Hanwell_WSSSPE} reflect on the 15-year history of open
source software development at \toolname{Kitware}.  In particular, they focus on their
success at growing their community of users through enabling multiple channels
of communication, directly reaching out to individuals, and lowering the
barrier to entry for contributions. This involves providing clear,
test-oriented and review-based mechanisms for evaluating contributions,
permissive licenses, and a service-based model for sustaining development.
This model enables \toolname{Kitware} to review both public funding, as well as private
funding to support improvements and targeted developments of the software.

Foster et al.~\cite{Foster_WSSSPE} discussed the approach of developing sustainability models
around Software-as-a-Service (SaaS) platforms, with the target example being
that of \toolname{GlobusOnline}.  The authors build a case that both grant-based and
volunteer-based development fall short in sustaining software, resulting in
software that is disproportionately difficult to use compared to its
functionality, which they note directly impacts the overall scientific
productivity of its users.  In contrast, by charging a subscription fee for
hosted, centrally-managed software (similar to offerings by \toolname{Dropbox}, \toolname{EverNote},
\toolname{GMail}), the authors propose to manage the funding cycle and enable a greater
focus on the aspects of software that directly impact individuals, rather than
funders.  \toolname{Globus} has deployed such a service, for which they have attempted to
develop a sustainable economic model that reduces institutional obstacles to
funding and subscription.  However, they do identify that cultural obstacles
do still remain, and they note a particular difference in culture between NSF-
and NIH-funded researchers.

While many papers briefly discussed licensing issues, Elster~\cite{Elster_WSSSPE}
directly discussed the impact of different licenses on obtaining industrial
funding for scientific software projects.  In particular, she describes
experiences with researchers unwilling to utilize GPL (copyleft) code, as it
adds restrictions to reuse that some industries find unacceptable for future
commercialization.  (This was discussed by Hanwell et al.~\cite{Hanwell_WSSSPE} as well.)
Further obstacles to sustainability identified by the author include patenting
of software, a lack of open access, and even more broadly, a lack of open
source codes even in open access journals.  Finally, a lack of awareness on the
part of scientific software developers of commodity libraries for common tasks
reduces their ability to reuse code.

\subsection{Education \& Training}

\todo{text about papers and discussion in this part of the workshop}

%\section{Other Discussion} \label{sec:other}
%
%\note{from the misc questions/answers/observations part of the google doc?}
%
%\note{all to put notes here of stuff that doesn't fit elsewhere, if any}
%
%\todo{perhaps remove this section}

\section{Cross-cutting Issues} \label{sec:cross-cutting}

Two issues were raised in a number of the workshops sessions, defining sustainability and career paths.  This section brings those discussions together.

\subsection{Defining Sustainability \note{lead: Daniel S. Katz}}  \label{sec:defining-sustainability}

%\note{So far this reads a lot like a collection of thoughts only, without ``red line''.}
%\note{Dan: assuming the previous note is from FL, what do you propose?}
The question of what was meant by ``sustainability'' in the context of software
came up in many different parts of the workshop, specifically in the first
keynote (\S\ref{sec:keynote1}), the Developing and Maintaining Software panel,
and the Policy panel.

In the opening keynote, Philip Bourne suggested that perhaps sustainability can
be defined as the effort that happens to make the essential things continue.
This leads to having to decide what it is that we want to sustain, whether what
we want to sustain is valuable, and finally, who would care if it went away,
and how one measures how much they care.

In the Developing and Maintaining Software panel, there was some discussion on
this topic: what does sustainability mean? It was pointed out that OSS Watch
proposes a Software Sustainability Maturity Model to address the issue of how
sustainability a particular element of software is, and says that this
sustainability is important. ``When choosing software for procurement or
development reuse - regardless of the license and development model you will
use - you need to consider the future. While a software product may satisfy
today's needs, will it satisfy tomorrow's needs? Will the supplier still be
around in five years' time? Will the supplier still care for all its customers
in five years' time? Will the supplier be responsive to bug reports and feature
requests? In other words, is the software sustainable?''~\cite{OSS-ssmm-web}

Attendees suggested that a key question that the definition of sustainability
is one issue on which the community needs to agree, and that ideally, an
initial definition would be determined during the workshop, or at least some
progress would be made towards this goal.  Another topic that came up is what
the goal of sustainability is.  Perhaps it is reproducible science, or
persistence, or quality, or something else.  Similarly, some attendees want to
understand how success in sustainability is measured.  How does a group of
developers know when they have actually achieved sustainable software? This led
to a comment that sustainability should be addressed throughout the full
software life cycle.

Another topic that came up during the Developing and Maintaining Software panel
is the relationship of sustainability to other software attributes.  Attendees
asked ``what is the relationship between sustainability and provenance?'' And,
``is usability separate from sustainability or a fundamental part of it?''

In addition, the Policy panel had a large amount of discussion about defining
sustainability, as one of the subtopics in that panel was Modeling
Sustainability, and modeling requires defining what will be modeled.

In this subtopic, two papers included discussion of the definition of
sustainability.  Venters et al.~\cite{Venters_WSSSPE} mentioned that this is a
rather ambiguous concept, and that the lack of an accepted definition gets in
the way of integrating the concept into software engineering. They suggest that
sustainability is a non-functional requirement, and that the quality of
software architectures determines sustainability. They then propose that
sustainability is a consequence of a set of central quality attributes:
extensibility, interoperability, maintainability, portability, reusability, and
scalability. Finally, they develop an architecture evaluation framework based
on scenarios that help to illuminate how to measure quality or sustainability
at the architectural level.

Katz and Proctor~\cite{Katz_WSSSPE} included a set of questions that could be
used to measure software sustainability, and though these questions might
falsely lead to yes or no answers, the complete set would determine a range of
values for sustainability. These questions are:
\begin{itemize}
\item Will the software continue to provide the same functionality in the future,
      even if the environment (other parts of the infrastructure) changes?
\item Is the functionality and usability clearly explained to new users?
\item Do users have a mechanism to ask questions and to learn about the software?
\item Will the functionality be correct in the future, even if the environment changes?
\item Does it provide the functionality that existing and future users want?
\item Does it incorporate new science/theory/tools as they develop?
\end{itemize}

Lenhardt~\cite{lenhardt-wssspe1-panel} summarized the contributions of the
Modeling Sustainability papers in the panel. As shown in
Table~\ref{tab:defining-sustainability}, for each paper he discussed what
software meant, whether there was a definition of sustainability, and what the
approach was to either understand or evaluate sustainability.

\begin{table}[t]
  \begin{scriptsize}
    \begin{center}
      \caption{Summary of Modeling Sustainability papers from Policy Panel.  Adapted from~\cite{lenhardt-wssspe1-panel}.}
      \label{tab:defining-sustainability}
      \begin{tabular}{|p{2.3cm}|p{3.6cm}|p{4.4cm}|p{4.8cm}|}
                \hline
{\bf Paper/Authors}
& {\bf Software}
& {\bf Sustainability}
& {\bf Approach to Understand or Evaluate Sustainability} \\
                \hline
Calero, et al.~\cite{Calero_WSSSPE}
& General notion of software. Not explicitly defined.
& Sustainability is linked to quality.
& Add to ISO \\
                \hline
Venters, et al.~\cite{Venters_WSSSPE}
& Software as science software; increasingly complex; service-oriented computing
& Extensibility, interoperability, maintainability, portability, reusability, scalability, efficiency
& Use various architecture evaluation approaches to assess sustainability \\
                \hline
Pierce, et al.~\cite{Pierce_WSSSPE}
& Cyberinfrastructure software
& Sustainable to the extent to which there is a community to support it
& Open community governance \\
                \hline
Katz and Proctor~\cite{Katz_WSSSPE}
& E-research infrastructures (i.e. cyberinfrastructure)
& Persisting over time, meeting original needs and projected needs
& Equates models for the creation of software with sustaining software \\
                \hline
Lenhardt, et al.~\cite{Lenhardt_WSSSPE}
& Broadly defined as software supporting science
& Re-use; reproducible science
& Comparing data management life cycle to software development life cycle \\
                \hline
Weber, et al.~\cite{Weber_WSSSPE}
& Software broadly defined; a software ecosystem
& Software niches
& Ecological analysis and ecosystem \\
                \hline
     \end{tabular}
    \end{center}
  \end{scriptsize}
\end{table}

Finally, in the workshop's closing session, one of the discussion topics was
what success would look like for the set of WSSSPE activities beyond just the workshop.
One of the answers that was suggested was
having an agreed version of what we mean by sustainability.

\subsection{Career Tracks for Scientific Software Developers} \note{lead: Daniel S. Katz} \label{sec:career-tracks}

Career track issues also came up repeatedly, starting in the the first keynote (\S\ref{sec:keynote1}), where Phil Bourne used the term ``the Google Bus'' to describe the issue of staff in his lab in San Diego who had chosen to leave science and work for Google (but since they didn't want to physically move, they take the Google bus to the Google office each day.)

The lack of career paths also came up in the Developing and Supporting Software discussion (\S\ref{sec:devel}), as an issue that led to difficulties in established projects retaining as well as recruiting staff. The issue there was that for software to be sustainable, there need to be career paths for software developers.  There was also some discussion about how should software development be balanced with research, with the idea that faculty are hired based on their research qualifications, not on their software development qualification, and as such, specializing in software development was likely to not lead to a career in science at a university, though clearly, national laboratories are a different case.

In the Policy discussion (\S\ref{sec:policy}), career paths were not discussed explicitly, though the discussion about credit, citation, and impact focused on how those who create or contribute to software are recognized; this recognition (or the current lack thereof) is tied to career paths (or their lack) for software developers.

The Communities discussion (\S\ref{sec:community}) did not directly address career paths, though one of the papers~\cite{Loffler_WSSSPE} that was part of this section also makes a similar policy point, in a section about career paths. Specifically, it says ``The most severe problem for developers in most computational sciences currently is that while most of the work is done creating hopefully well-written, sustainable software, the academic success is often exclusively tied to the solution of the scientific problem the software was designed for. Tasks that from a software engineering standpoint are essential, e.g., high usability, well-written and updated documentation, or porting infrastructure to new platforms, are not rewarded within this system.''

In terms of possible solutions, the Developing and Supporting Software discussion said that things are beginning to improve, for example, with incentives for software development in the form of altmetrics, tenure committee consideration, and NSF's change in its biosketch requirement: asking for proposer's ``products'' rather than their ``publications,'' as had been done previously.
However, it was also noted it can be very difficult to measure where people are using small bits of your code.  And the possible career paths for a specialist software developer working as part of a scientific research group in a university are certainly not yet clear.

\section{Case Studies \note{lead: Ketan Maheshwari, contributors: Andy Terrel}} \label{sec:use-cases}

% == Notes and ToDos ==
% 1. Why do we classify software cases into general purpose and special
% purpose? what is different? What are the sustainability challenges each face?
% Are they radically different? What kind of research do each category support?
% Are there separate questions to ask and answer for each category? Can we
% assume that the general purpose software has a wider scope and user community
% and challenges specific to these kinds of environments are not relavent in
% the special purpose scientific software which tend to be developed in the
% small?

% 2. What are the factors that has driven the development of these cases? How
% was the development started? What was the spark? Was there a design approach
% followed? Is the software modular/extensible? Can we identify unique and
% common factors and the reasons behind them.

% 3. Can we characterize the process itself for the selected cases:
% collaborative, separate.

% 4. ToDo: summarize and report on an inclusive list (might not be exhaustive)
% of all the software products discussed in the workshop. (If possible color
% code the representation according to a scheme based on the different
% dimensions pertaining to them.)

% 5. Has one or more of the following best practices played a role in the development: 
%     a) hosted community source-code and other artefacts repository
%     b) documentation, wikis and training material
%     c) mailing lists and web-presence
%     d) development process: agile, hackathons, collaborative?
%     e) regression tests, continuous integration, quality dashboard, code review

% 6. Is there a conscious effort to deal with issues such as backward
% compatibility, ...

% 7. Career path: are the original developer(s) still around? Is most of the
% discussion around career path issues a myth after all?

% 8. Is there a software paper for this software product? Cite?

% 9. Has the project proved to be sustainable so far? Does the developers,
% managers make conscious efforts to make sure the project is sustainable? What
% determines the sustainability of the project in question: science,
% publications, usability, funding, reusability? Perhaps all factors to some
% extent? What factors are dominating? Can we send out a message to current
% scientific project to emphasize on one or more of these or other aspects?
% ====

In this section, we discuss some of the software projects as case studies to
better understand the points discussed during the workshop and described in the
previous sections. We dive deeper to find how are they affected by the
sustainability related issues in practice. Additionally, we discuss issues
surrounding people, especially the career paths of original developers and
investigate to answer the question: \emph{are the original developers still
around?}

We classify the software projects discussed in the workshop in two broad
categories. First, the \emph{utility} software, comprising of general purpose
software. Utility software is often used as enabler and/or facilitator for the
development of other tools and techniques to carry out scientific work. This
includes the software developed to efficiently utilize new research
infrastructures. Second, the \emph{scientific} software, comprising the
software that was originally developed with an aim to solve a specific
scientific problem.

% Why are we classifying software like above?
This classification is motivated by our argument that the two kinds of software
projects wildly vary in factors such as scope, purpose and usage. The
development and management of each kind is significantly different.
Consequently, the sustainability challenges faced by them differ and must be
treated separately. For instance, the challenges faced by a gateway software
development project such as \toolname{CIPRES} or a visualization software product
such as \toolname{VisIT} are distinct to a niche software for \textit{ab initio}
modeling and simulation such as \toolname{VASP} or \toolname{Quantum Espresso}.

\subsection{Utility Software}
Software developed with a potentially wider audience and general purpose usage
in mind is utility software. Utility software typically does not address
fundamental research problems for a given scientific domain. Examples are
collaborative development frameworks such as \toolname{GitHub} and
\toolname{Bitbucket}, distributed workflow and generic computing frameworks
such as \toolname{Galaxy}, \toolname{nanoHUB}, \toolname{SimGrid}, \toolname{Swift},
\toolname{Globus} and \toolname{VisTrails}, and visualization frameworks such as
\toolname{VisIT}.

% What are the sustainability challenges? What are the opportunities?
Development is often a high risk/reward undertaking exploring uncharted
territories and is largely influenced by (re)usability factors. Owing to
relatively large number of features, the development and prototype process is
also lengthy which poses a significant survival risk. Challenges on a class of
utility software for new architectures is well discussed
in~\cite{Ferenbaugh_WSSSPE}.

On the other hand, utility software presents opportunities to be usable by a
larger community making its undertaking and development an attractive pursuit.
It is generally more visible in community which in turn leads to a broader and
deeper participation. For instance, promoting collaborations across the breadth
(e.g.  different departments) and depth (e.g. stakeholders within a department)
of community, one of the key ingredient of a sustainable process. Successful
utility projects reap high rewards and have a longer usage span. Development
process becomes user-driven and self-sustaining.

One such example is the \toolname{Galaxy} project~\cite{Galaxy}. It follows agile software
development practices and implements standards such as test-driven development
and socialized bug managing practices via \toolname{trello}. Galaxy
\emph{histories} and \emph{toolshed} offers easy community sharing of data and
tools further promoting a collaborative environment. The project very closely
follows the guidelines described in Carver and Thiruvathukal~\cite{Carver_WSSSPE} and many
from Prli\'{c} and Procter~\cite{Prlic_WSSSPE}. Many utility software projects are often developed
aiming better utilizing a particular, new infrastructure and architecture,
e.g., \toolname{MVAPICH}, \toolname{VisIT}. Similarly, to leverage the power of
emerging architectures such as accelerators, new code and libraries are
required. The experience of one such effort as described
in Ferenbaugh~\cite{Ferenbaugh_WSSSPE} which met with a limited success but nonetheless
with many invaluable lessons were learned about influential cultural and
technical aspects in sustainable software development practices.

A relatively new paradigm in utility software is the software delivered as
service over the web. With increasing popularity of cloud-based storage and
computational environments, many users are leaning towards tools used as
services. \toolname{GitHub} and \toolname{Bitbucket} can be argued to be
such tools, catering to collaborative development. For scientific users
\toolname{Globus}-based tools are a case of service-based utility software
discussed during the workshop. The data movement and sharing services offered
by Globus can be easily used over the web by collaborating researchers.
%
%
%\note{Dan: so is the only difference here that this category is aimed at new
%systems, where utility software is not?  If so, perhaps this should be
%incorporated into that subsection, so there would just be two subsections.
%Something in the framing of this subsection doesn't feel right to me.  FL: I
%respectfully disagree the the only reason to develop frameworks is to better
%utilize (new or old) infrastructure. Another really important reason is to
%modularize development, making it much easier for an individual developer/user
%to integrate their own code. Ketan: I think in this section I wanted to bring
%up a class of software projects which are undertaken with the main aim of
%tackling new systems and architectures in mind. New infrastructures such as
%GPU, Cloud, fast IB, federal science networks, etc. were on my mind as I see
%many submissions around them in the workshop.}
%
%

\subsection{Scientific software}
Scientific software consists primarily of special-purpose software that was
purpose-built for a target use-case scenario,  fixed/frozen requirements in
mind or solving a specific problem. Software projects pertaining to specific
scientific domain tend to be in a niche and the user community tends to be
small to medium. They are mostly driven by the science and specific needs of a
research group. Specific needs such as numerical accuracy and
algorithmic optimization are some of the paramount requirements of most
scientific software.

% What are the challenges?
Long-term sustainability of scientific software is often a significant
challenge and face radically different issues compared to utility software.
Many submissions reported that software is practically considered a ``byproduct''
of the actual research. Others contended that the software was not the main
funded part of their research. A smaller codebase and fixed requirements result
in stability, ease of installation, and configurability.  Many such projects
mature and are treated as libraries to be used into larger systems such as some
of the utility software discussed in the previous section. While the software
can stay stable and require relatively low maintenance, the responsibility is
often on the shoulders of a very few developers who might not be specialists in
software development. Development tends to be linear and simplistic with a
limited scope to follow software best practices.

Some examples of such software discussed as part of the workshop are,
\toolname{DUNE}, \toolname{R/qtl}, \toolname{Kitware}, \toolname{PETSc}, \toolname{MINRES-QLP},
\toolname{FASTMath}.  Most of the above efforts are focused on one scientific
and/or applied mathematics domain. However, sometimes such projects grow beyond
the initial vision of developers. One such example is \toolname{Kitware}, which
while being a software product specializing in the scientific process, has a
core focus of developing communities around software processes. An instance of
this process is the development of the \toolname{CMake} build utility, which
started out as a building tool for \toolname{ITK} but grew to become a generic
build utility for C++ projects. Similarly, \toolname{PETSc} is growing towards
becoming a general purpose utility system usable for solving variety of
scientific problems.

Finally, it appears that the career options of developers from different kinds
of software projects differ significantly. In order to further investigate the
much discussed issue of career paths for developers of these software, we
conducted a small survey of the use-cases discussed above with respect to the
current status of its core developers. We looked at the workshop submissions
and the original cited papers for the respective software. Where available, we
looked at the webpages and source code of the earliest versions to find the
original developers. We included projects older than 5 years and excluded
relatively new projects such as \toolname{SciPy} and \toolname{numFOCUS} from this
survey. we find that in a majority of scientific software more than one
original developers are still around, in some cases for more than 10 years,
(eg. \toolname{R/qtl}).  Whereas, in the case of utility software projects, it
appears that at least one original developer is still around, often in a
part-time or ``benefactor'' role. In other cases, it appears that the original
developers have went on to assume a larger role, with the same project.

In conclusion, we find that there are distinctions in the characteristics and
the kinds of challenges faced between utility and scientific software projects.
In terms of career paths of original developers, it appears that the career
choices for utility software developers are broader compared to the developers
of scientific software. However, it appears that in the case of scientific
software, core competencies of original developers tend to be domain science
who happen to be the coders of their software. As projects as a whole, the
small-scale, specific software have a lower risk and cost in terms of
sustaining as compared to utility software. We also find that the utility
software tend to adhere more loyally to the many best practices discussed
during the workshop. Often the sustainability of scientific software projects
is achieved by the fact that the core developer or team heavily utilizes the
software for their science, e.g.,  \toolname{R/qtl, PETSc}.  Furthermore, the
development of scientific software requires more scientific skills compared
with those of utility software. This diminishes the need of specialist
programmers and practices in many cases where the bulk of development is
carried out by a domain scientist. With these inferences, we believe that
separate guidelines and/or principles for sustainability be defined and
practiced for each category of scientific software and they be given separate
treatment.

\note{Dan: I think there's some overlap
  between this and the discussion in the Developing and Support
  Software section.  Perhaps this should go in the cross-cutting
  issues section?  Let's write it down here - then we can decide if it goes elsewhere}
\note{Ketan: yes, may be this section will split and merge into other sections}

\section{Conclusions} \label{sec:conclusions}

\todo{pull the discussion together}

\todo{add some analysis of the deficiencies and difficulties that are
  present in different fields, and those that are common?}

\todo{say something about licensing - lessons, advice, etc.?}

\subsection{Top Issues}

\subsection{Recommendations or Lessons}

\note{if needed.}

\subsection{Follow up actions}

\note{or at least the discussion about them, and the current plans for
  future events.}

\note{conclusions from pre-workshop paper follows}

The WSSSPE workshop has begun an experiment in how we can
collaboratively build a workshop agenda. However, contributors also
want to get credit for their participation in the process. And the
workshop organizers want to make sure that the workshop content and
their efforts are recorded.  Ideally, there would be a service that
would be able to index the contributions to the workshop, serving the
authors, the organizers, and the larger community. But since there
isn't such a service today, the workshop organizers are writing this
initial report and making use of arXiv as a partial solution to
provide a record of the workshop.

After the workshop, one or more additional papers will be created that
will include the discussions at the workshop. These papers will likely
have many authors, and may be submitted to peer-reviewed journals.

\katznote{discuss sustainability survey}

\section*{Acknowledgments}

\todo{anyone who needs to put something in here should}

Some of the work by Katz was
supported by the National Science Foundation while working at the
Foundation; any opinion, finding, and conclusions or recommendations
expressed in this material are those of the author(s) and do not
necessarily reflect the views of the National Science Foundation. HL
was supported by the National Evolutionary Synthesis Center (NESCent), 
NSF \#EF-0905606.

Chue-Hong and Hetherington were supported by the UK Engineering and 
Physical Sciences Research Council (EPSRC) Grant EP/H043160/1 for the 
UK Software Sustainability Institute.

\appendix
\section{Papers Accepted and Discussed at WSSSPE1} \label{sec:papers}

\section*{Developing and Supporting Software}

\subsubsection*{Development Experiences}

\begin{itemize}

\item Mark C. Miller, Lori Diachin, Satish Balay, Lois Curfman
  McInnes, Barry Smith. Package Management Practices Essential for
  Interoperability: Lessons Learned and Strategies Developed for
  FASTMath~\cite{Miller_WSSSPE}

\item Karl W. Broman, Thirteen years of R/qtl: Just barely sustainable~\cite{Broman_WSSSPE}

\item Charles R. Ferenbaugh, Experiments in Sustainable Software
  Practices for Future Architectures~\cite{Ferenbaugh_WSSSPE}

\item Eric G Stephan, Todd O Elsethagen, Kerstin Kleese van Dam, Laura
  Riihimaki. What Comes First, the OWL or the Bean?~\cite{Stephan_WSSSPE}

\item Derek R. Gaston, John Peterson, Cody J. Permann, David Andrs,
  Andrew E. Slaughter, Jason M. Miller, Continuous Integration for
  Concurrent Computational Framework and Application Development~\cite{Gaston_WSSSPE}

\item Anshu Dubey, B. Van Straalen. Experiences from Software
  Engineering of Large Scale AMR Multiphysics Code Frameworks~\cite{Dubey_WSSSPE}

\item Markus Blatt. DUNE as an Example of Sustainable Open Source
  Scientific Software Development~\cite{Blatt_WSSSPE}

\item David Koop, Juliana Freiere, Cl\'{a}udio T. Silva, Enabling
  Reproducible Science with VisTrails~\cite{Koop_WSSSPE}

\item Sean Ahern, Eric Brugger, Brad Whitlock, Jeremy S. Meredith,
  Kathleen Biagas, Mark C. Miller, Hank Childs, VisIt: Experiences
  with Sustainable Software~\cite{Ahern_WSSSPE}

\item Sou-Cheng (Terrya) Choi. MINRES-QLP Pack and Reliable
  Reproducible Research via Staunch Scientific Software~\cite{Choi_WSSSPE}

\item Michael Crusoe, C. Titus Brown. Walking the talk: adopting and
  adapting sustainable scientific software development processes in a
  small biology lab~\cite{Crusoe_WSSSPE}

\item Dhabaleswar K. Panda, Karen Tomko, Karl Schulz, Amitava Majumdar.
The MVAPICH Project: Evolution and Sustainability of an Open Source
Production Quality MPI Library for HPC~\cite{Panda_WSSSPE}

\item Eric M. Heien, Todd L. Miller, Becky Gietzel, Louise
  H. Kellogg. Experiences with Automated Build and Test for
  Geodynamics Simulation Codes~\cite{Heien_WSSSPE}

\end{itemize}

\subsubsection*{Deployment, Support, and Maintenance of Existing Software}

\begin{itemize}

\item Henri Casanova, Arnaud Giersch, Arnaud Legrand, Martin Quinson,
  Fr\'{e}d\'{e}ric Suter. SimGrid: a Sustained Effort for the
  Versatile Simulation of Large Scale Distributed
  Systems~\cite{Casanova_WSSSPE}

\item Erik Trainer, Chalalai Chaihirunkarn, James Herbsleb. The Big
  Effects of Short-term Efforts: A Catalyst for Community Engagement
  in Scientific Software~\cite{Trainer_WSSSPE}

\item Jeremy Cohen, Chris Cantwell, Neil Chue Hong, David Moxey,
  Malcolm Illingworth, Andrew Turner, John Darlington, Spencer
  Sherwin. Simplifying the Development, Use and Sustainability of HPC
  Software~\cite{Cohen_WSSSPE}

\item Jaroslaw Slawinski, Vaidy Sunderam. Towards Semi-Automatic
  Deployment of Scientific and Engineering Applications~\cite{Slawinski_WSSSPE}

\end{itemize}

\subsubsection*{Best Practices, Challenges, and Recommendations}

\begin{itemize}

\item Andreas Prli\'{c}, James B. Procter. Ten Simple Rules for the
  Open Development of Scientific Software~\cite{Prlic_WSSSPE}

\item Anshu Dubey, S. Brandt, R. Brower, M. Giles, P. Hovland,
  D. Q. Lamb, F. Löffler, B. Norris, B. O'Shea, C. Rebbi, M. Snir,
  R. Thakur, Software Abstractions and Methodologies for HPC
  Simulation Codes on Future Architectures~\cite{Dubey2_WSSSPE}

\item Jeffrey Carver, George K. Thiruvathukal. Software Engineering
  Need Not Be Difficult~\cite{Carver_WSSSPE}

\item Craig A. Stewart, Julie Wernert, Eric A. Wernert, William
  K. Barnett, Von Welch. Initial Findings from a Study of Best
  Practices and Models for Cyberinfrastructure Software Sustainability~\cite{Stewart_WSSSPE}

\item Jed Brown, Matthew Knepley, Barry Smith. Run-time extensibility:
  anything less is unsustainable~\cite{Brown_WSSSPE}

\item Shel Swenson, Yogesh Simmhan, Viktor Prasanna, Manish Parashar,
  Jason Riedy, David Bader, Richard Vuduc. Sustainable Software
  Development for Next-Gen Sequencing (NGS) Bioinformatics on Emerging
  Platforms~\cite{Swenson_WSSSPE}

\end{itemize}

\section*{Policy}

\subsubsection*{Modeling Sustainability}

\begin{itemize}

\item Coral Calero, M. Angeles Moraga, Manuel F. Bertoa. Towards a
  Software Product Sustainability Model~\cite{Calero_WSSSPE}

\item Colin C. Venters, Lydia Lau, Michael K. Griffiths, Violeta
  Holmes, Rupert R. Ward, Jie Xu. The Blind Men and the Elephant:
  Towards a Software Sustainability Architectural Evaluation Framework~\cite{Venters_WSSSPE}

\item Marlon Pierce, Suresh Marru, Chris Mattmann. Sustainable
  Cyberinfrastructure Software Through Open Governance~\cite{Pierce_WSSSPE}

\item Daniel S. Katz, David Proctor. A Framework for Discussing
  e-Research Infrastructure Sustainability~\cite{Katz_WSSSPE}

\item Christopher Lenhardt, Stanley Ahalt, Brian Blanton, Laura
  Christopherson, Ray Idaszak. Data Management Lifecycle and Software
  Lifecycle Management in the Context of Conducting Science~\cite{Lenhardt_WSSSPE}

\item Nicholas Weber, Andrea Thomer, Michael Twidale. Niche Modeling:
  Ecological Metaphors for Sustainable Software in Science~\cite{Weber_WSSSPE}

\end{itemize}

\subsubsection*{Credit, Citation, Impact}

\begin{itemize}

\item Matthew Knepley, Jed Brown, Lois Curfman McInnes, Barry
  Smith. Accurately Citing Software and Algorithms Used in
  Publications~\cite{Knepley_WSSSPE}

\item Jason Priem, Heather Piwowar. Toward a comprehensive impact
  report for every software project~\cite{Priem_WSSSPE}

\item Daniel S. Katz. Citation and Attribution of Digital Products:
  Social and Technological Concerns~\cite{Katz2_WSSSPE}

\item Neil Chue Hong, Brian Hole, Samuel Moore. Software Papers:
  improving the reusability and sustainability of scientific software~\cite{Chue_Hong_WSSSPE}

\end{itemize}

In addition, the following paper from another area were also discussed
in this area.

\begin{itemize}

\item Frank L\"{o}ffler, Steven R. Brandt, Gabrielle Allen and Erik
  Schnetter. Cactus: Issues for Sustainable Simulation Software~\cite{Loffler_WSSSPE}

\end{itemize}

\subsubsection*{Reproducibility}

\begin{itemize}

\item Sou-Cheng (Terrya) Choi. MINRES-QLP Pack and Reliable
  Reproducible Research via Staunch Scientific Software~\cite{Choi_WSSSPE}

\item Victoria Stodden, Sheila Miguez. Best Practices for
  Computational Science: Software Infrastructure and Environments for
  Reproducible and Extensible Research~\cite{Stodden_WSSSPE}

\end{itemize}

\subsubsection*{Implementing Policy}

\begin{itemize}

\item Randy Heiland, Betsy Thomas, Von Welch, Craig Jackson. Toward a
  Research Software Security Maturity Model~\cite{Heiland_WSSSPE}

\item Brian Blanton, Chris Lenhardt, A User Perspective on Sustainable
  Scientific Software~\cite{Blanton_WSSSPE}

\item Daisie Huang, Hilmar Lapp. Software Engineering as
  Instrumentation for the Long Tail of Scientific Software~\cite{Huang_WSSSPE}

\item Chandra Krintz, Hiranya Jayathilaka, Stratos
  Dimopoulos, Alexander Pucher, Rich Wolski. Developing Systems for API Governance~\cite{Krintz_WSSSPE}

\end{itemize}

\section*{Communities, Models, and Education}

\subsubsection*{Communities}

\begin{itemize}

\item Reagan Moore. Extensible Generic Data Management Software~\cite{Moore_WSSSPE}

\item Karen Cranston, Todd Vision, Brian O'Meara, Hilmar Lapp. A
  grassroots approach to software sustainability~\cite{Cranston_WSSSPE}

\item J.-L. Vay, C. G. R. Geddes, A. Koniges, A. Friedman,
  D. P. Grote, D. L. Bruhwiler. White Paper on DOE-HEP Accelerator
  Modeling Science Activities~\cite{Vay_WSSSPE}

\item Ketan Maheshwari, David Kelly, Scott J. Krieder, Justin M. Wozniak, Daniel S. Katz, Mei Zhi-Gang, Mainak Mookherjee. Reusability in Science: From Initial User Engagement to Dissemination of Results~\cite{Maheshwari_WSSSPE}

\item Edmund Hart, Carl Boettiger, Karthik Ram, Scott Chamberlain. rOpenSci -- a collaborative effort to develop R-based tools for facilitating Open Science~\cite{Hart_WSSSPE}

\item L. Christopherson, R. Idaszak, S. Ahalt. Developing Scientific Software through the Open Community Engagement Process~\cite{Christopherson_WSSSPE}

\item Marlon Pierce, Suresh Marru, Mark A. Miller, Amit Majumdar, Borries Demeler. Science Gateway Operational Sustainability: Adopting a Platform-as-a-Service Approach~\cite{Pierce2_WSSSPE}

\item Lynn Zentner, Michael Zentner, Victoria Farnsworth, Michael
  McLennan, Krishna Madhavan, and Gerhard Klimeck, nanoHUB.org:
  Experiences and Challenges in Software Sustainability for a Large
  Scientific Community~\cite{Zentner_WSSSPE}

\item Andy Terrel. Sustaining the Python Scientific Software Community~\cite{Terrel_WSSSPE}

\item Frank L\"{o}ffler, Steven R. Brandt, Gabrielle Allen and Erik
  Schnetter. Cactus: Issues for Sustainable Simulation Software~\cite{Loffler_WSSSPE}

\item Nancy Wilkins-Diehr, Katherine Lawrence, Linda Hayden, Marlon Pierce, Suresh Marru, Michael McLennan, Michael Zentner, Rion Dooley, Dan Stanzione. Science Gateways and the Importance of Sustainability~\cite{Wilkins-Diehr_WSSSPE}

\item Sou-Cheng (Terrya) Choi. MINRES-QLP Pack and Reliable
  Reproducible Research via Staunch Scientific Software~\cite{Choi_WSSSPE}

\end{itemize}

In addition, the following paper from another area was also discussed
in this area.

\begin{itemize}

\item Marcus Hanwell, Amitha Perera, Wes Turner, Patrick O'Leary,
  Katie Osterdahl, Bill Hoffman, Will Schroeder. Sustainable Software
  Ecosystems for Open Science~\cite{Hanwell_WSSSPE}

\end{itemize}

\subsubsection*{Industry \& Economic Models}

\begin{itemize}

\item Anne C. Elster. Software for Science: Some Personal Reflections~\cite{Elster_WSSSPE}

\item Ian Foster, Vas Vasiliadis, Steven Tuecke. Software as a Service
  as a path to software sustainability~\cite{Foster_WSSSPE}

\item Marcus Hanwell, Amitha Perera, Wes Turner, Patrick O'Leary,
  Katie Osterdahl, Bill Hoffman, Will Schroeder. Sustainable Software
  Ecosystems for Open Science~\cite{Hanwell_WSSSPE}

\end{itemize}

In addition, the following papers from other areas were also discussed
in this area.

\begin{itemize}

\item Brian Blanton, Chris Lenhardt, A User Perspective on Sustainable
  Scientific Software~\cite{Blanton_WSSSPE}

\item Markus Blatt. DUNE as an Example of Sustainable Open Source
  Scientific Software Development~\cite{Blatt_WSSSPE}

\item Dhabaleswar K. Panda, Karen Tomko, Karl Schulz, Amitava
  Majumdar. The MVAPICH Project: Evolution and Sustainability of an
  Open Source Production Quality MPI Library for HPC~\cite{Panda_WSSSPE}

\item Andy Terrel. Sustaining the Python Scientific Software Community~\cite{Terrel_WSSSPE}

\end{itemize}

\subsubsection*{Education \& Training}

\begin{itemize}

\item Ivan Girotto, Axel Kohlmeyer, David Grellscheid, Shawn
  T. Brown. Advanced Techniques for Scientific Programming and
  Collaborative Development of Open Source Software Packages at the
  International Centre for Theoretical Physics (ICTP)~\cite{Girotto_WSSSPE}

\item Thomas Crawford. On the Development of Sustainable Software for
  Computational Chemistry~\cite{Crawford_WSSSPE}

\end{itemize}

In addition, the following papers from other areas were also discussed
in this area.

\begin{itemize}

\item Charles R. Ferenbaugh, Experiments in Sustainable Software
  Practices for Future Architectures~\cite{Ferenbaugh_WSSSPE}

\item David Koop, Juliana Freiere, Cl\'{a}udio T. Silva, Enabling
  Reproducible Science with VisTrails~\cite{Koop_WSSSPE}

\item Sean Ahern, Eric Brugger, Brad Whitlock, Jeremy S. Meredith,
  Kathleen Biagas, Mark C. Miller, Hank Childs, VisIt: Experiences
  with Sustainable Software~\cite{Ahern_WSSSPE}

\item Sou-Cheng (Terrya) Choi. MINRES-QLP Pack and Reliable
  Reproducible Research via Staunch Scientific Software~\cite{Choi_WSSSPE}

\item Frank L\"{o}ffler, Steven R. Brandt, Gabrielle Allen and Erik
  Schnetter. Cactus: Issues for Sustainable Simulation Software~\cite{Loffler_WSSSPE}

\item Erik Trainer, Chalalai Chaihirunkarn, James Herbsleb. The Big
  Effects of Short-term Efforts: A Catalyst for Community Engagement
  in Scientific Software~\cite{Trainer_WSSSPE}

\end{itemize}

\section{Attendees \note{lead: Shel Swenson}} \label{sec:attendees}

The following is a partial list of attendees who were recorded on the
Google doc~\cite{WSSSPE1-google-notes} that was being used for live note taking at the workshop, or by the SC13 student volunteers, with some additions also made by the authors of this report.

\begin{multicols}{3}
\setlength{\parindent}{0pt}
%in theory, should save the old value then set it back after this section, but since this is the end...

Jay Alameda

Gabrielle Allen

David Andrs

Brian Austin

Lorena A. Barba

David Bernholdt

Phil Bourne

Karl Broman

Sharon Broude Geva

Jed Brown

Maxine Brown

David Bruhwiler

Bruno Bzeznik

Alexandru Calotoiu

Jeffrey Carver

Shreyas Cholia

Peng Huat Chua

Neil Chue Hong %surname is Chue Hong, this is the correct order.

John W. Cobb

Timothy Cockerill

Karen Cranston

Rion Dooley

Anshu Dubey

Marat Dukhan

Ian Foster

Juliana Freire

Jeffrey Frey

Derek Gaston

Allison Gehrke

Brian Glendenning

Christian Godenschwager

Derek Groen

Edmund Hart

Magne Haveraaen

Steven Heaton

Oscar Hernandez

James Hetherington

Simon Hettrick

Jonathan Hicks

Kenneth Hoste

James Howison

Daisie Huang

Shao-Ching Huang

Tsutomu Ikegami

Kaxuya Ishimura

Christian Iwainsky

Craig Jackson

Wesley Jones

Randall Judd

Shusuke Kasamatsu

Daniel S. Katz

Kerk Kee

Kellie Kercher

Mads Kristensen

Carolyn Lauzon

Arnaud Legrand

Chris Lenhardt

Michael Levy

Frank L\"{o}ffler

Monica L\"{u}cke

Simon A. F. Lund

Arthur Maccabe

Paul Madden

Louis Maddox

Philip Maechling

Ketan Maheshwari

Brian Marker

Suresh Marru

Cezary Mazurek

James McClure

Matt McKenzie

Chris Mentzel

Paul Messina

Mike Mikailov

J. Yates Monteith

Reagan More

Rafael Morizawa

Pierre Neyron

Lucas Nussbaum

Patrick O'Leary

Manish Parashar

Cody Permann

Jack Perdue

John Peterson

Quan Pham

Marlon Pierce

Heather Piwowar

David Proctor

Sara Rambacher

Nicolas Renon

Jason Riedy

Todd Rimer

Bill Sacks

Andreas Schreiber

William Scullin

Andrew Slaughter

Jaraslaw Slawinski

Arfon Smith

Spencer Smith

James Spencer

Eric Stahlberg

Timothy Stitt

Hyoshin Sung

Fr\'{e}d\'{e}ric Suter

Shel Swenson

Yoshio Tanaka

Andy Terrel

George  Thiruvathukal

Keren Tomko

John Towns

Erik Trainer

Satori Tsuzuki

Matthew Turk

Eric van Wyk

Colin C. Venters

Brice Videau

Tajendra Vir Singh

Von Welch

Nancy Wilkins-Diehr

Theresa Windus

Felix Wolf

Rich Wolski

Lynn Zentner

\end{multicols}

\bibliographystyle{vancouver}

\bibliography{wssspe_paper}
\end{document}