### 2015-09-30 WSSSPE3/report/wg_appendix_template.tex
Rename wg_appendix_software_credit.tex to wg_appendix_template.tex

\section{Software Credit Working Group Discussion}

\subsection{Group Members}
{\small
\begin{longtable}{ll}
   Alice Allen           &  Astrophysics Source Code Library (ASCL)
\\ Daniel S. Katz        &  University of Chicago \& Argonne National Laboratory
\end{longtable}
}

\subsection{Summary of Discussion}

\subsection{Description of Opportunity, Challenges, and Obstacles}

\subsection{Key Next Steps}

\subsection{Plan for Future Organization}

\subsection{What Else is Needed?}

\subsection{Key Milestones and Responsible Parties}

\subsection{Description of Funding Needed}