### 2015-09-30 WSSSPE3/report/wg_main_legacy_SW.tex
For main report
Former-commit-id: a5b6ce61a25ea1be2db5109d7f96e0cbf386a995

\subsection{Topic} \todo{change the title to your topic}

\subsubsection{Why it is important}
\todo{short text here}

\subsubsection{Fit with related activities}
\todo{short text here - can include links/cites}

\subsubsection{Discussion}
\todo{short-ish text here}

\subsubsection{Plans}
\todo{short text here - not bullets}

\subsubsection{Landing Page}
\todo{link to landing page}