### 2015-10-09 WSSSPE3/report/wg_appendix_software_credit.tex
SC: added some discussion to appendix, other subsections completed
Former-commit-id: 2b2bb8d4d76e58b73579a1e60e21905b9c2c381b

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Software Credit Working Group Discussion}
\label{sec:appendix_SW_credit}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Kyle Niemeyer\footnote{email: \href{mailto:kyle.niemeyer@oregonstate.edu}{kyle.niemeyer@oregonstate.edu}} will serve as the point of contact for this working group, and be responsible for ensuring timely progress of the planned actions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Group Members}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{itemize}
\item Alice Allen -- Astrophysics Source Code Library
\item Sou-Cheng Choi -- NORC at University of Chicago, Illinois Institute of Technology
\item James Hetherington -- University College London
\item Lorraine Hwang -- University of California, Davis
\item Daniel S.\ Katz -- University of Chicago, Argonne National Laboratory
\item Frank Löffler -- Louisiana State University
\item Abby Cabunoc Mayes -- Mozilla Science Lab
\item Kyle E.\ Niemeyer -- Oregon State University
\item Grace Peng -- National Center for Atmospheric Research
\item Ilian Todorov -- Science \& Technology Facilities Council, UK
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Summary of Discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The following section summarizes the working group's discussion based on contributions prior to the meeting~\cite{WSSSPE3-SC-github-issues} and the collaborative notes taken during the meeting~\cite{WSSSPE3-SC-google-notes}.
Please refer to the original sources for the unedited discussions.

Initial discussions focused on both various mechanisms for, and the philosophical approach behind, crediting software in scientific papers.
These began with proposals for various ways to credit software (or other research products including data) that contributed more significantly than a generic citation, including
\begin{itemize}
    \item A hierarchy of citations, with a ``substantial'' citation category to indicate software or data that played a more significant role in the research;
    \item Transitive credit~\cite{wssspe2_katz,Katz:2014_tc}, which assigns contriponents (contributors and components) various weights depending on their level of importance; and
    \item Project CRediT~\cite{projectcredit}, which assigns roles to paper authors based on their specific contributions; and
    \item Mozilla Science Lab's recently introduced Contributorship Badges for Science, which provide a badge---associated with ORCID~\cite{orcid}---based on author contributions in a similar manner to Project CRediT.
\end{itemize}
However, as of this writing, only the Project CRediT roles~\cite{McCall2015_credit,Lin2015_credit} and Contributorship Badges~\cite{} have been implemented in published papers, though both of these only provide a single ``Software'' or ``Computation'' category associated with software.
In addition, neither of these options allow for the citation of software itself, but only provide an author contribution related to software.
The discussion quickly focused on transtitive credit as a more quantitative measure of allocating credit to both authors and software, although there were some concerns about authors overestimating their own contributions compared to prior work.

The discussion then evolved into philosophical questions about the importantance or reliance of a particular work on prior science, materials, or software---in other words, is there a difference between depending on prior scientific advances and depending on particular software (or experimental equipment)?
Multiple contributors converged on the conclusion that unique capabilities require some additional credit.
The---albeit limited---consensus was that if a particular study relied on the unique capabilities of software, data, or an experimental apparatus, then the authors or developers that created this capability should be credited somehow.

The group also agreed on the fact that additional data was required to support the assertion that software was not being sufficiently cited in the literature.
In particular, this issue seemed to be field-dependent.
For example, as shown by study of Howison and Bullard~\cite{Howison2015}, in the field of biology the most-cited papers appear to be describing scientific software.
However, this may not---and likely is not---the case in other fields, nor is it clear whether developers of scientific software, even in the case of the biology field, are receiving sufficient credit for their efforts.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Description of Opportunity, Challenges, and Obstacles}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

There currently is no standard/accepted mechanism for citing software or receiving credit for software (akin to publications/citations).
Software is eligible for DOI assignment, but DOI metadata fields are not well tuned or standardized for software (vs. publications).
Some software providers apply for DOIs but it is still not widely adopted.
Also, there is no mechanism to cite software dependencies within software.

Obstacle: indexers (e.g., Scopus, Web of Science, Google Scholar) don't currently support software/data document types or DataCite DOIs.

Obstacle: no standard mechanism for software to cite other software.

Obstacle: for tenure, there is no standard for software products to be included in policy,
sometimes not even within one university: hard to change globally.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Key Next Steps}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{enumerate}
\item Hold virtual meeting to determine group members responsible\slash willing to work on the following tasks, to be organized within one month of the workshop
\item Compile best practices of software citation across multiple disciplines, including journals and communities of interest\slash practice in the research world, to begin by December 2015
\item Compile examples of including other products in promotion and tenure dossier, to begin by December 2015
\item Draft the Software Citation Principles document (including citation metadata file), by April 2016
\item Publish\slash release the Software Citation Principles document, by August 2016
\item Reach out to journals, publishers, teachers\slash educators, indexers, and professional societies---likely through meetings with key groups, to begin by September 2016

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Plan for Future Organization}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The WSSSPE breakout group plans to join efforts related to citing software with the FORCE11 Software Citation Working Group (FORCE11-SCWG)\footnote{FORCE11 Software Citation Working Group, \url{https://www.force11.org/group/software-citation-working-group}}; Kyle Niemeyer formally requested the merging of these groups following the meeting.
However, some future plans of the WSSSPE group fall outside the scope of FORCE11-SCWG, which covers software citation practices.
These activities include working with indexers such as Web of Science and Scopus to index software citations archived on, e.g., Zenodo\footnote{Zenodo, \url{https://zenodo.org}}, and pursuing the development of an open indexing service; such plans will be pursued either separately or through the formation of follow-on FORCE11 working groups.

The group will primarily communicate electronically, with Kyle Niemeyer responsible for ensuring regular progress.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{What Else is Needed?}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The near-term actions of the group, focused mainly on software citation, do not require any additional resources.
However, connections with publishers and indexers will be needed to pursue related activities, although the FORCE11-SCWG may satisfy this need; in addition, some members of the group already reached out to relevant contacts.
Funding may be needed to organize meetings or for group members to attend relevant meetings, as discussed further below.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Key Milestones and Responsible Parties}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Following the meeting, Kyle Niemeyer formally requested the merging of software citation activities with FORCE11-SCWG.
Within a month of the meeting, Kyle will organize a virtual meeting of the group and lead the division of responsibilities for compiling existing practices of software citation and including software\slash products in promotion and tenure dossiers.
Building off of these efforts, the next major milestone is drafting the Software Citation Principles document in collaboration with the SCWG, targeted for April 2016.
While the existing directors of the SCWG, Arfon Smith and Dan Katz, lead the efforts of that group towards the Software Citation Principles document, Kyle will help coordinate contributions from the WSSSPE group members.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Description of Funding Needed}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Some funding would be useful to support primarily travel to conferences for group meetings (e.g., FORCE2016\footnote{FORCE2016, \url{https://www.force11.org/meetings/force2016}}), and to hold meetings to bring together both group members and key stakeholders (e.g., journals, publishers, professional societies, indexers).
In addition, funding would be desired to support group members' time to perform work towards the key steps described previously.