### 2015-10-22 WSSSPE3/report/wg_main_software_credit.tex
Added SC landing pages

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Software Credit Working Group}
\label{sec:software-credit}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Why it is important}
\todo{short text here}

\subsubsection{Fit with related activities}
\todo{short text here - can include links/cites}

Publishing Software Working Group (\S\ref{sec:publishing-software}): publishing a software paper offers one mechanism for receiving credit

\subsubsection{Discussion}
\todo{short-ish text here}

\subsubsection{Plans}
\todo{short text here - not bullets}

\subsubsection{Landing Page}

Since near-term efforts will be shifting to the FORCE11 Software Citation Working Group (SCWG), we direct interested readers to that group's existing landing page\footnote{FORCE11-SCWG landing page, \url{https://www.force11.org/group/software-citation-working-group}} and GitHub repository\footnote{FORCE11-SCWG GitHub page,  \url{https://github.com/force11/force11-scwg}}.