### 2015-10-27 WSSSPE3/report/wg_appendix_software_credit.tex
Finished draft of credit appendix

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Software Credit Working Group Discussion}
\label{sec:appendix_SW_credit}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Kyle Niemeyer\footnote{email: \href{mailto:kyle.niemeyer@oregonstate.edu}{kyle.niemeyer@oregonstate.edu}} will serve as the point of contact for this working group, and be responsible for ensuring timely progress of the planned actions.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Group Members}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{itemize}
\item Alice Allen -- Astrophysics Source Code Library
\item Sou-Cheng Choi -- NORC at University of Chicago, Illinois Institute of Technology
\item James Hetherington -- University College London
\item Lorraine Hwang -- University of California, Davis
\item Daniel S.\ Katz -- University of Chicago, Argonne National Laboratory
\item Frank Löffler -- Louisiana State University
\item Abby Cabunoc Mayes -- Mozilla Science Lab
\item Kyle E.\ Niemeyer -- Oregon State University
\item Grace Peng -- National Center for Atmospheric Research
\item Ilian Todorov -- Science \& Technology Facilities Council, UK
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Summary of Discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The following section summarizes the working group's discussion based on contributions prior to the meeting~\cite{WSSSPE3-SC-github-issues} and the collaborative notes taken during the meeting~\cite{WSSSPE3-SC-google-notes}.
Please refer to the original sources for the unedited discussions.

Initial discussions focused on both various mechanisms for, and the philosophical approach behind, crediting software in scientific papers.
These began with proposals for various ways to credit software (or other research products including data) that contributed more significantly than a generic citation, including
\begin{itemize}
    \item A hierarchy of citations, with a ``substantial'' citation category to indicate software or data that played a more significant role in the research;
    \item Transitive credit~\cite{wssspe2_katz,Katz:2014_tc}, which assigns contriponents (contributors and components) various weights depending on their level of importance; and
    \item Project CRediT~\cite{projectcredit}, which assigns roles to paper authors based on their specific contributions; and
    \item Mozilla Science Lab's recently introduced Contributorship Badges for Science~\cite{Mozilla_badges}, which provide a badge---associated with ORCID~\cite{orcid}---based on author contributions in a similar manner to Project CRediT.
\end{itemize}
However, as of this writing, only the Project CRediT roles~\cite{McCall2015_credit,Lin2015_credit} and Contributorship Badges~\cite{Mozilla_badges} have been implemented in published papers, though both of these only provide a single ``Software'' or ``Computation'' category associated with software.
In addition, neither of these options allow for the citation of software itself, but only provide an author contribution related to software.
The discussion quickly focused on transtitive credit as a more quantitative measure of allocating credit to both authors and software, although there were some concerns about authors overestimating their own contributions compared to prior work.

The discussion then evolved into philosophical questions about the importantance or reliance of a particular work on prior science, materials, or software---in other words, is there a difference between depending on prior scientific advances and depending on particular software (or experimental equipment)?
Multiple contributors converged on the conclusion that unique capabilities require some additional credit.
The---albeit limited---consensus was that if a particular study relied on the unique capabilities of software, data, or an experimental apparatus, then the authors or developers that created this capability should be credited somehow.

The group also agreed on the fact that additional data was required to support the assertion that software was not being sufficiently cited in the literature.
In particular, this issue seemed to be field-dependent.
For example, as shown by study of Howison and Bullard~\cite{Howison2015}, in the field of biology the most-cited papers appear to be those describing scientific software.
However, this may not---and likely is not---the case in other fields, nor is it clear whether developers of scientific software, even in the case of the biology field, are receiving sufficient credit for their efforts.

During the first day of WSSSPE3 breakout sessions, the group discussed the Entertainment Identifier Registry (EIDR)\footnote{Entertainment Identifier Registry: \url{http://eidr.org}} as a potential model for scientific software.
That system assigns unique Digital Object Identifiers (DOIs)---the same system used for scientific publications---to all content (e.g., movies, television shows) and contributors, along with relevant metadata.
One important use of this system is to track rights and credit for contributors to entertainment works in order to distrubute revenues---similar to the proposed transitive credit concept.

The group also discussed separating quantitative measures (e.g., number of citations) from the value of a work in order to give credit, moving towards qualitative or anecdotal evidence of value.
Other topics brought up included a form of PageRank~\cite{Brin1998} for citations, based on number of mentions, and using market penetration or adoption rate in a community as a metric, although it was not clear how this would be measured.
Finally, the concept a software tool's uniqueness or indispensability to a community was mentioned, with value being characterized by a particular piece of software either offering unique capabilities or doing something better, faster, or with less computational requirements than other offerings.

On the second day of WSSSPE3, the working group decided to put aside the taxonomy of contributions and focus on software citations to ensure developers receive credit (regardless of contribution).
Eventually, once standardizing software citations, the goal would be to return to establishing different roles\slash contributions for this credit.
Following this decision, the group identified two necessary actions to move forward:
\begin{enumerate}
    \item standardizing a citation file or some other form of metadata associated with software, and
    \item standardizing the way to cite software (used directly) in papers.
\end{enumerate}
For both of these actions, a number of ongoing efforts were identified and discussed.

\subsubsection{Software Citation Metadata}

At a minimum, the metadata needed for software citation (aside from the DOI for the software itself) includes:
\begin{itemize}
    \item Contributors, in the form of name and ORCID,
    \item Software dependencies, in the form of DOI, and
    \item Other people and artifacts that would be cited or acknowledged in a paper.
\end{itemize}
This information would then be contained in a citation file as part of 
the, e.g., GitHub repository.
The group also discussed similar efforts such as Codemeta\footnote{Codemeta: \url{https://github.com/codemeta/codemeta}}, an attempt to codify minimal 
metadata schemes in JSON and XML for scientific software and code, and 
implementing transitive credit via JSON-LD~\cite{wssspe2_katz}.
Some questions arose about how this information would be stored for 
closed-source software.

As one mechanism for constructing accurate contributor lists from existing 
project contributors, the group discussed associating GitHub accounts---as 
well as accounts on Bitbucket, CodePlex, and other repositories for 
open-source scientific software---accounts with ORCID.
However, a (quick) response from GitHub (via Arfon Smith) indicated that 
this might not be possible in the near future: ``GitHub doesn't have any 
plans to allow ORCID accounts to be associated with GitHub user accounts.''

\subsubsection{Citing Software in Publications}

Although far from a standard practice, examples of citing software in 
publications can be found in various scientific communities---notably, 
representative samples can be found in astronomy\footnote{Astronomy software citation examples: \url{https://docs.google.com/document/d/1q9ULl7alA3veL7Qwg7jGteRWeJwlrkvRHSXjvt-rTs0}} and biology~\cite{Howison2015}.
The group recommended collecting similar examples from other communities, then 
developing a software citation principles document in concert with the 
FORCE11 Software Citation Working Group (see \S\ref{SC:plan} for more details) 
following the model of the FORCE11 Data Citation Principles 
document~\cite{DataCitation2014}.

The group also discussioned briefly whether software used directly in 
a publication---whether to perform simulation, analysis, or as a dependency 
for new software developed---should be distinguished from other references 
due to the dependence of the study on these research artifacts.
Suggestions included a separate list of citations (with DOIs) for software 
and other research objects that serve this sort of ``vital'' role.
Similar recommendations were made by the credit breakout group at 
WSSSPE2~\cite{WSSSPE2}.

Finally, although a discrete task from the software citation, significant 
discussion focused on ensuring software citations are indexed in the same 
manner as publications---allowing the construction of a corresponding 
software citation network.
Currently, software releases can receive citable DOIs via 
Zenodo\footnote{Zenodo, \url{https://zenodo.org}} and 
FigShare\footnote{FigShare: \url{https://figshare.com}}; however, these 
citations are not processed by indexers such as Web of Science, Scopus, 
or Google Scholar.
Thus, either in parallel or following the primary task, the group will 
need to reach out to these organizations 
Initial conversations with Elsevier\slash Scopus (via Michael Taylor) 
during WSSSPE3 clarified that Scopus is not yet DataCite DOI aware, and 
also does not yet have an internal identifier for software or data (but 
needs\slash plans to add this support).
According to Taylor, they prefer a ``software article'' with the usual 
article metadata (e.g., authors, citations), and mentioned Zenodo as an 
example---this proposal seemed to align with group discussions.
Taylor also mentioned another benefit of the software and associated DOI 
on GitHub: in addition to a citation, could obtain statistics on 
usage/downloads/forks.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Description of Opportunity, Challenges, and Obstacles}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Modern scientific research relies considerably on software, and ensuring 
that developers of such scientific software receive credit for their efforts 
will encourage additional creation and maintenance.
There currently is no standard mechanism for citing software or receiving credit for software (akin to publications/citations).
Software is eligible for DOI assignment, but DOI metadata fields are not well tuned or standardized for software (vs. publications).
Some software providers apply for DOIs but it is still not widely adopted.
Also, there is no mechanism to cite software dependencies within software.

Major obstacles include the fact that indexers (e.g., Scopus, Web of Science, 
Google Scholar) do not currently support software/data document types or 
DataCite DOIs.
Therefore, even with univeral association of scientific software with DOIs 
and standardizing practices for citing software in publications

Although the working group discussions at WSSSPE3 did not focus much on the 
topic of tenure, the group recognized that there is no standard policy---even 
within a single university---for software products to be included in 
promotion and tenure dossiers.
Thus, it may be difficult to encourage valuing software contributions across 
the United States or United Kingdom, and globally; furthermore, stakeholders 
are typically not tenured and thus may not be influential enough.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Key Next Steps}
\label{SC:next-steps}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{enumerate}
\item Hold virtual meeting to determine group members responsible\slash willing to work on the following tasks, to be organized within one month of the workshop
\item Compile best practices of software citation across multiple disciplines, including journals and communities of interest\slash practice in the research world, to begin by December 2015
\item Compile examples of including other products in promotion and tenure dossier, to begin by December 2015
\item Draft the Software Citation Principles document (including citation metadata file), by April 2016
\item Publish\slash release the Software Citation Principles document, by August 2016
\item Reach out to journals, publishers, teachers\slash educators, indexers, and professional societies---likely through meetings with key groups, to begin by September 2016

\end{enumerate}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Plan for Future Organization}
\label{SC:plan}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The WSSSPE breakout group plans to join efforts related to citing software with the FORCE11 Software Citation Working Group (FORCE11-SCWG)\footnote{FORCE11 Software Citation Working Group, \url{https://www.force11.org/group/software-citation-working-group}}; Kyle Niemeyer formally requested the merging of these groups following the meeting.
However, some future plans of the WSSSPE group fall outside the scope of FORCE11-SCWG, which covers software citation practices.
These activities include working with indexers such as Web of Science and Scopus to index software citations archived on, e.g., Zenodo or FigShare, and pursuing the development of an open indexing service; such plans will be pursued either separately or through the formation of follow-on FORCE11 working groups.

The group will primarily communicate electronically, with Kyle Niemeyer responsible for ensuring regular progress.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{What Else is Needed?}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The near-term actions of the group, focused mainly on software citation, do not require any additional resources.
However, connections with publishers and indexers will be needed to pursue related activities, although the FORCE11-SCWG may satisfy this need; in addition, some members of the group already reached out to relevant contacts.
Funding may be needed to organize meetings or for group members to attend relevant meetings, as discussed further below.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Key Milestones and Responsible Parties}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Following the meeting, Kyle Niemeyer formally requested the merging of software citation activities with FORCE11-SCWG.
Within a month of the meeting, Kyle will organize a virtual meeting of the group and lead the division of responsibilities for compiling existing practices of software citation and including software\slash products in promotion and tenure dossiers.
Building off of these efforts, the next major milestone is drafting the Software Citation Principles document in collaboration with the SCWG, targeted for April 2016.
While the existing directors of the SCWG, Arfon Smith and Dan Katz, lead the efforts of that group towards the Software Citation Principles document, Kyle will help coordinate contributions from the WSSSPE group members.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Description of Funding Needed}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Some funding would be useful to support primarily travel to conferences for group meetings (e.g., FORCE2016\footnote{FORCE2016, \url{https://www.force11.org/meetings/force2016}}), and to hold meetings to bring together both group members and key stakeholders (e.g., journals, publishers, professional societies, indexers).
In addition, funding would be desired to support group members' time to perform work towards the key steps described previously.