### 2015-10-28 WSSSPE3/report/wg_main_software_credit.tex
draft SC main section completed

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Software Credit Working Group}
\label{sec:software-credit}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsubsection{Why it is important}

Science and engineering research increasingly relies on software, but there is currently no standard mechanism for citing software or receiving credit for developing software akin to receiving credit via citations for writing papers.
Standardizing software citations offers one route to establishing such a mechanism.
Software is currently eligible for DOI assignment, but DOI metadata fields are not well tuned for software compared to publications. 
Some software providers apply for DOIs but it is still not widely adopted. 
Also, there is no mechanism to cite software dependencies within software in the same way papers cite supporting prior work.

\subsubsection{Fit with related activities}

Publishing Software Working Group (\S\ref{sec:publishing-software}): publishing a software paper offers one existing mechanism for receiving credit, and further developing new publishing concepts for software will strengthen our activities.

A number of groups external to WSSSPE (although with some overlapping members) are also focused on aspects of software credit, including the FORCE11 Software Citation Working Group (see plans for coordination below).
In addition, a Software Credit workshop\footnote{London Software Credit workshop: \url{www.software.ac.uk/software-credit}} convened in London on October 19, following the conclusion of WSSSPE3.
See \S\ref{sec:appendix_SW_credit} for more detailed discussion of related activities.

\subsubsection{Discussion}

The group discussed a number of topics related to software credit, including a contributorship taxonomy, software citation metadata, standards for citing software in publications, and increasing the value of software in academic promotion and tenure reviews.
Although initial discussions both prior to and during WSSSPE3 focused on contribution taxonomy and dividing credit, discussing as an example the Entertainment Identifier Registry\footnote{Entertainment Identifier Registry: \url{http://eidr.org}} used in the entertainment industry, the group decided to prioritize software citation.
This decision was motivated by the idea that standardizing citations for software would introduce some initial credit for developers, and later the quantification of credit could be refined based on concepts such as transitive credit~\cite{wssspe2_katz,Katz:2014_tc}.

The majority of the remaining discussion focused on standardizing (1) the metadata necessary for software to be cited and (2) the mechanism for citing software in publications.
In addition, discussions also oriented around the indexing of software citations necessary for establishing a software citation network either integrated with the existing paper citation ecosystem or complementary to it.
See \S\ref{sec:appendix_SW_credit} for a more detailed summary of the working group's discussion on these topics.

\subsubsection{Plans}

The group already merged with the FORCE11 Software Citation Working Group (SCWG), and their efforts will focus (over the next six to nine months) on developing a document describing principles for software citation.
Following publication of that document, the working group will focus on outreach to key groups (e.g., journals, publishers, indexers, professional societies).
Longer-term plans include working with indexers to ensure that software citations are indexed and pursuing an open\slash community indexer; these activities may be organized by future FORCE11 working groups.

\subsubsection{Landing Page}

Since near-term efforts will be shifting to the FORCE11-SCWG, we direct interested readers to that group's existing landing page\footnote{FORCE11-SCWG landing page, \url{https://www.force11.org/group/software-citation-working-group}} and GitHub repository\footnote{FORCE11-SCWG GitHub page,  \url{https://github.com/force11/force11-scwg}}.