### 2015-10-29 WSSSPE3/report/wg_appendix_publishing_SW.tex
Some formatting fixes to SW pub appendix

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Publishing Software Working Group Discussion}
\label{sec:appendix_publishing_SW}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Steven R.\ Brandt\footnote{email: \href{mailto:sbrandt@cct.lsu.edu}{sbrandt@cct.lsu.edu}} will serve as the point of contact for this working group.

\subsection{Group Members}

\begin{itemize}
\item Steven R. Brandt -- Louisiana State University
\item Daniel Gunter -- LBNL
\item Yuhan Ding -- Illinois Institute of Technology
\item Neil Chue Hong -- Software Sustainability Institute
\end{itemize}

\subsection{Summary of Discussion}

A tentative first cut at the list containing executable papers identified the following:
\begin{itemize}
\item ACM Transactions on Mathematical Software (TOMS): provides the extra step
 of having reviewers validate the code which was submitted with the publication.
\item The Mathematica Journal: publishes Mathematica notebooks (with equations,
figures, etc.) directly.
\item O'Reily Media: announced that it plans to make IPython Notebooks a first-class
 authoring environment for their publishing program alongside their existing mechanisms.
\item Nature: offering a list of notebooks published alongside more traditional articles,
 and is looking at ways to make these documents more official. There are, in fact, a
 number of journals that offer ``electronic supplements'' to the more traditionally published
 static articles.
\item IPython: maintained list of ``reproducible academic publications''~\cite{ipython-pubs}.
\item KBase: narratives build on IPython\slash Jupyter notebooks to build publications that are
  reproducible, and can be commented or annotated.
\end{itemize}

The group also discussed future possibilities for this new publication format which might
provide advantages:
\begin{itemize}
\item Journals could be built around an existing, widely used framework thereby reducing
  the burden of studying code on the part of reviewers (common bits of infrastructure
  which aren't relevant to the science would be automatically excluded).
\item Journals might be encouraged to use more metadata, making them easier to mine
  for various analytical purposes.
\item RIO is an effort to publish fragmentary results that can subsequently be combined
  into a single content item.
\item Papers could be made more understandable. Each equation or technical term could
  offer a link to a document/tutorial explaining its origin and/or derivation.
\item So many options for publication currently exist that good science may be getting
  lost in the noise. Would some sort of ``upvote'' mechanism be of value?
\item Some sort of Replicated Computation Results badge could be made available to
  publications that have undergone greater scrutiny (this is already done by TOMS).
\end{itemize}

\subsection{Description of Opportunity, Challenges, and Obstacles}

The opportunity is to collect a list of current executable papers and
shine a light on the experiments and development efforts currently underway.

The only obstacle to this is the difficulty in finding and identifying such
publications. The Software Sustainability Institute was able to do something similar
for publications about software by making a public page containing a catalog
of these publications and enlisting the help of the community to grow the list.

\subsection{Key Next Steps}

Create the first version of the web page to be displayed on the Software Sustainability
Institute's website.

\subsection{Plan for Future Organization}

None at this time.

\subsection{What Else is Needed?}

Nothing else at this time.

\subsection{Key Milestones and Responsible Parties}

Steven R. Brandt will create a first version of the page within a week or so of the WSSSPE3 conference.

Neil Chue Hong will take responsibility for the page once it is up.

\subsection{Description of Funding Needed}

None.