### 2015-10-30 WSSSPE3/report/wg_main_publishing_SW.tex
Fixed to do, and tweaked wording.

\subsection{Publishing Software Working Group Discussion} \label{sec:publishing-software}

\subsubsection{Why it is important}

This group explored the value of executable papers (papers whose content
includes the code needed to produce their own results), and other forms of
publishing which include dynamic electronic content. Transitioning to this type
of publication offers possibilities of addressing, or partially addressing
sustainability concerns such as reproducibility, software credit, and best
practices.

\subsubsection{Fit with related activities}

\textbf{Reproducibility}: Part of the purpose of these executable paper venues
is to (at least partially) address the reproducibility issue by making the
paper itself recompute its own results.

\textbf{Software Credit (\S\ref{sec:software-credit})}: Since these forms of
publishing must make their sources explicit in order to execute, they should be
easier to trace even if appropriately worded credit for software is not
provided. In addition, these notebooks make it possible to provide/define
additional metadata to make the tracing of credit clearer. Finally,
attributions could be added to citations to identify whether a paper extends a
result, verifies it, contradicts it, etc.

\textbf{Best Practices (\S\ref{sec:best-practices})}: Because an executable paper
showcases the code, and the code itself is subject to the review process,
authors are more likely to pay attention to coding practices. In addition,
because the paper must explain what the code does, better documentation is
likely to be achieved.

\subsubsection{Discussion}

The group felt that the best way to encourage the use of these new publishing
concepts would be to create and curate a list identifying publishing venues
that support them. The Software Sustainability Institute agreed to host this list.

\subsubsection{Plans}

The plan is to create a web page describing executable papers, their value, and
a list of what publishers support them.

\subsubsection{Landing Page}

The aforementioned page to be published on SSI.