### 2015-11-07 WSSSPE3/report/wg_main_user_community.tex
edits & suggestions in user community section

\subsection{Building Sustainable User Communities for Scientific Software}

%\subsubsection{Why it is important}

User communities are the lifeblood of sustainable scientific software. The user community includes the developers, 
both internal and external, of the software; direct users of the software; other software projects that depend on
the software; and any other groups that create or consume data that is specific to the software. Together these
groups provide both the reason for sustaining the software and, collectively, the requirements that drive its continued
evolution and improvement.

\subsubsection{Fit with related activities}

Mozilla Science maintains a ``Working Open Project Guide'' (\url{http://mozillascience.github.io/leadership-training}). \katznote{should get changed to a cite} From the introduction:
\begin{quote}
Working openly with contributors enables your
    community to learn how to build and collaborate together. This
    document is a guideline on how to work openly and involve others
    in your projects with Mozilla. We want to help you engage your
    community in a way that encourages contributors and builds other
    leaders.
 \end{quote}

Several books have been written about software communities:
\begin{itemize}
\item ``Art of Community'' by Jono Bacon. \katznote{cite?} We could consider distilling this for scientific software.
\item Iain Larmour, from EPSRC in the UK (EPSRC: https://www.epsrc.ac.uk/) (not sure who from mentioned UK Collaborative Computational Projects (CCP: http://ccp.ac.uk)
\end{itemize}

\subsubsection{Discussion}

Discussion revolved around a few questions: what is the benefit of having a ``community'' for software sustainability, what
practices and circumstances lead to having a community, how can funding help or hinder this process, and perhaps most
importantly how can best practices be described and distilled into a document that can help new projects.

The benefits of having a community that were brought up were considered largely obvious. In addition to having advocates for
the software, and a possible source of ``free'' contributions to the codebase, the community becomes a good source for
requirements, feedback, and metrics. The software community can also act as ``cheerleaders'' who convince funders or other
potential users to fund/use the software, and thus help sustain the software.

Practices and circumstances that lead to a community are first, that the software offers value. But in addition to this, a
community will be much more likely to form if they receive (expert) support when they have questions. Additional contributing
factors are good usability (not always needed), and an open development process such as IPython developer meetings on YouTube.
It was also pointed out that an evangelist for the project, not necessarily but often one of the developers, can often make a
big difference.

Funding can help the process by encouraging both value to the community and high-quality user support. Only providing funding
for the software development may create good software, but with less likelihood to have a real community. It was discussed
that federal laboratories are a good incubator for software communities, and that a general facility like EarthCube is too
dispersed to really make a community. Also, domain-specific groups within laboratories or universities might provide as an
incubator for software communities.

In describing best practices, the group discussed the different modes for starting a scientific software project: building on
an existing product that needs improving, recognizing an unsatisfied need of an existing community, or creating a new solution
to a need not yet recognized by the community. The group also thought that the existing books on software communities would
need to be evaluated in light of differences between Science Software projects and general OSS projects in terms of scale,
science, acknowledgement and credit, and funding models.

Additional details on the group's discussion can be found in Appendix~\ref{sec:appendix_user_community}.

\subsubsection{Plans}

The most important next steps is a ``Best Practice'' document, which would describe what successful projects with engaged
communities look like, how to replicate this type of project, and look at end-of-life on a community project. Inputs to this
document would include a software community survey of high functioning communities such as R Open Science, Python SciPy,
OPeNDAP, and Unidata, with analysis of factors that feed into their success. Also references lik the ``Art of Community'' could
be adapted and summarized for the science software community.

Another next step would be increasing recognition of need for science software projects to focus on building and supporting
their user community. Good Software Engineering practices are not enough, and popular training like Software Carpentry does
not currently address this issue head on.

\subsubsection{Landing Page}
\todo{link to landing page}