### 2015-11-09 WSSSPE3/report/wg_appendix_funding_spec_expert.tex
Fix typos

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Funding Specialist Expertise Group Discussion}
\label{sec:appendix_funding_spec_expert}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

James Hetherington\footnote{email: \href{mailto:j.hetherington@ucl.ac.uk}{j.hetherington@ucl.ac.uk}}
will serve as the point of contact for this working group, and be responsible for ensuring timely progress of the planned actions.

\subsection{Group Members}

The group at WSSPE:

\begin{itemize}
\item Don Middleton -- National Center for Atmospheric Research
\item Joshua Greenberg -- Alfred P. Sloan Foundation
\item James Hetherington -- University College London
\item Lindsay Powers -- The HDF Group
\item Mark A. Miller -- San Diego Supercomputer Center
\item Dan Sellars -- CANARIE
\end{itemize}

This was further enhanced by additional discussions at the following
GCE15 conference:

\begin{itemize}
\item Lorraine Hwang -- UC Davis
\item Simon Trigger
\item Nancy Wilkins-Diehr -- San Diego Supercomputer Center
\item Alexander Vyushkov -- Notre Dame
\item Sandra Gesing - Notre Dame
\item Ali Swanson -- University of Oxford

\end{itemize}

\subsection{Summary of Discussion}

In addition to the points noted in the main discussion~\ref{RSE}, we also
discussed the following:

``Are you an RSE or a RA?'' -- this is not properly a binary question. Most of
us sit at different points on that spectrum, and move along it during our
careers. (Ususally from RA to RSE -- examples of movement in the other direction
from readers would be welcomed.)
Either way, the label ``Research Software Engineer'' is now starting to
have some power. Many scientists do not want to be writing code; some do, to
varying degrees. These groups can usefully support each other.

he power of the label? Getting the word out about RSE support using the label.

Will research science developers be required, in the long run?
One issue that came up was whether the need for developers was a time bounded one; is it the case that the new generation of computer and software savvy scientists will be so comfortable in developing their own code that the professional developer will not be needed.  And this brings up the flip side question, “Do scientists really want to be writing code?”
Career Path.
We also had a little discussion about how to make a career path for research developers. It need not be solely an academic enterprise, but tenure is always problematic for people of this class.

Skills and resources may vary between teams. To help resolve this, maintaining
high levels of communication between groups will be valuable. In the UK, there
are plans to permit resource sharing between institutional RSE groups.

Collaborative funding can be important for RSE groups, to ensure that research
leadership remains with the domain scientists. At NCAR, University partnerships
are required for submission of proposals, so collaboration is an essential part
of grant submission, and this will tend to bring developers and
scientists together. The UCL group also follows this approach, with all bids
requiring an academic collaborator.
Domain scientists and developers are funded together in a single proposal.
Another example of a success is the development of semantics and linked data in
support of Ocean Sciences. An Earthcube funded project pairs domain scientists
with RSEs and has been successful as the semantics attached have increased data
use and discovery significantly.

An alternative approach has been the provision of programming expertise as
part of national compute services. The US XSEDE project's extended collaboration
 service (ECS) is a set of developers who are paid with XSEDE funding,
 and are on “permanent” staff.  When PIs request allocations on XSEDE resources,
 there is a finite pool of developer time that can be awarded,
typically for one year only, and at partial effort, typically 20 percent or so.
The finite time allowed provides motivation for the scientist and their group to
work closely the developer and to become educated in what the developer is doing,
so they can sustain the effort, once the ECS period is over. This funding
mechanism can be highly efficient for scientific problems, because the developer
pool assembled by the research providers are, by definition, expert in the
characteristics of their specific resource, and can very quickly assess the
scientists needs, and what it will take to implement software that meets the
user’s needs. However, it does not develop capacity within institutions,
and indeed can have a ``chilling effect''.

The UK allows this kind of collaboration to support the creation of scientific
software for the large supercomputing resource called ARCHER.
However, as well as resourcing this from the staff of the Edinburgh Parallel
Computing Centre, who host the computer, this ``embedded CSE'' resource also
allows the programming to come from local groups. This has been very helpful
in providing funding to establish local groups. These work best when they
develop good collaborations with national cyberinfrastructure pools.
When an organization assembles
a developer pool, diversity is developed and skills can be transferred.

We would like to see these models applied outside High Performance Computing;
most scientific software is not destined to run on national cyberinfrastructure,
but needs similar support. The argument regarding making better use of expensive
hardware through software improvements has been useful politically, (and many
RSE groups are cited in organizations which host clusters for this reason), but
the time has come to make the case that software itself is a critical
cyberinfrastructure, and, with a much longer shelf-life than hardware, is itself
a capital investment.

The Canarie group (Canada) accepts proposals
for providing services to broad communities, integrate people who are doing
things that are complementary, the goal is to make the available stack more
robust and richer for everyone. They offer short rounds of funding that can
have as a key metric creating some useful functionality that shows a diversity
of input and draws from across disciplines, then more funding could follow. This
could apply within or across institutions.

There can be problems communicating across cultural barriers, with domain
scientists seeing developers as “other”.
Collaboration, and tools to fund, encourage or motivate collaboration are
extremely important.

We think support from non-governmental organizations will be important if RSE
groups will become established.
The Sloan Foundation is currently funding data science engineers, who work in
the context of other SW developers at University of Washington.
These scientists work in the e-Science Studio/dataScience Studio, and they
help a group of graduate students in solving their problems in data
science and data management. During Fall and Spring, a 10-week incubator program
allows students to work two days a week to work on a data-intensive science
project. Some fraction of the developer time is dedicated to the developers'
personal interests as well as instruction.

The goal for Sloan is to provide success stories, to provide demonstrable value
in the presence of data scientists on university staff. These stories are the
basis for arguments to the host organization. This is an effort to create
awareness of the value of research scientist developers. Embedding with
scientists, and adding spare capacity is critical to making the innovation
possible. This model is essentially to argue for permanent budget lines to
support data scientists as part of university staff hires, just as with core
facilities. This could become a fee-for-service model requested by grant
funding, just as DNA sequencing is for core facilities, if it becomes apparent
that this gives competitive advantage to the University’s research effort.

One model that has been helpful in finding funding for RSE groups is the use of
funds left over on research grants when RAs have left prematurely -- PIs like
this arrangement as it is hard to find good staff for short-term positions, so
having a pool of research programming staff on hand resolves this problem.
We recommend that funders give explicit guidance to grant holders and
institutions that such arrangement are favorable. Framework agreements
permitting this to go ahead without
checking back every time with funders and/or grant panels would further
smooth this.
(This is also provides a more stable job for those who hold these skills, but
arguments about making life nicer for postdocs will not help persuade funders
or PIs!)

There is some question about the most effective
duration and percent of full time for a programmer's work on a project.
At least three months is necessary for the programmer to read into the science
(RSEs must not become so disengaged from research that they don't have time to
read a few papers -- this will result in code which doesn't meet scientific
needs), but too long could result in an RSE losing their flexibility, becoming
so engaged in one project that when that project ends, they find it hard to
transfer. For this reason, we also recommend that 40 percent is ideal;
two projects per developer, with some time for training and infrastructure work.
Two developers per project seems to be ideal, in the sense that software
development is enhanced by two pairs of eyes.

There is as yet no clear answer as to the scale of aggregation needed to
make such a program work. A university wide program allows enough scale to
be robust to fluctuations of funding within one field.
But a specialization focus on developers to support, for example,
physical or biological
sciences may be preferable, if the customer base is large enough.
The desire to aggregate enough work to make it sustainable, and the need to
have domain-relevant research programming skills, are in tension.

In the UK, another source of funding for research software has been the
Collaborative computational projects (CCPs): domain specific communities put
forward proposals that are a priority of the community as a whole, for example,
biosimulation or plasma physics. These bodies act as custodians of community codes,
and a central team also provides software engineering support.

However this area develops, the need for funding to secure software as a
cyberinfrastructure component is clear. Funding which permits code to be
refactored, tidied and optimized  is rare; this is often done ``on the sly''
in a scientifically focused grant. The UK EPSRC's ``software for the future''
call, which really permits explicit investment in software as an infrastructure,
is so oversubscribed as to have a FOUR PERCENT success rate; the demand is clear!

One opportunity is the idea of co-design; where infrastructural libraries are
developed alongside the scientific codes that will call them. However,
collaboration is hard to foster here; as incentive structures are still focused
on short-term papers. This can cause infrastructure developers to
focus more on publications
in their areas of mathematics and CS, the domain developers on the shorter-term
needs of their own fields. Genuine collaborative co-construction
is harder to foster.

It can be a more difficult
to help leading domain scientists see the value of engineering effort than those
in their teams who are forced to work with difficult-to-use or unreliable
software tools, as they do not see the pain. Perhaps a version of
``software carpentry'' targeted at those PIs
who are awarded or apply for software heavy grants could be valuable here.

RSEs provide a useful contribution to their Universities' *teaching* missions,
as well as for research, as they are well placed to deliver the research programming
training that many scientists now need. In the
longer term, with programming skills taught to all through their career,
we hope specialist
scientific developers will be less needed.

\subsection{Key Next Steps}

We will seek to identify and approach existing research programming organizations,
to get their permission to list them on a list of research software groups.
We will also look for examples of groups which have successfully become self-
sustaining following

\subsection{Plan for Future Organization and Future Needs}

The UKRSE community will provide initial facilities to host this list, and
continue to work to spread the initiative, but local leadership in the US
is needed if this campaign is to succeed. This will require an initial gathering
of identified research software organizations in the US to this end.

\subsection{Description of Funding Needed}

Financial support for an initial conference to bring together research software
groups to form an organization and create a resource sharing structure would
help to further this campaign.

In the longer term, funding organizations, especially non-governmental organizations
with the capability to effect innovation through seed funding, could provide
support to nucleate the creation of research software groups.