### 2015-11-23 WSSSPE3/report/wg_appendix_best_practices.tex
Appendix D: Fix typos.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Best Practices Group Discussion}
\label{sec:appendix_best_practices}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\todo{add POC here}

\subsection{Group Members}
\katznote{add affils for all, please}

\begin{itemize}
\item Abani Patra, University at Buffalo
\item Sandra Gesing, University of Notre Dame
\item Neil Chue Hong, Software Sustainability Institute
\item Gregory Tucker, University of Colorado at Boulder
\item Birgit Penzens, California State University Long Beach
\item Abigail Cabunoc Mayes, Mozilla Foundation
\item Jeff Carver, University of Alabama
\item Frank L\"{o}ffler, Louisiana State University 
\item Colin Venter,  University of Huddersfield
\item Lorraine Hwang, UC Davis - CIG
\item Sou-Cheng Choi, NORC at the University of Chicago \&  Illinois Institute of Technology
\item Suresh Marru, Indiana University
\item Don Middleton, NCAR 
\item Daniel Katz,  U Chicago \& Argonne
\item Kyle Niemeyer, Oregon State University
\item Jeffrey Carver, University of Alabama
\item Dan Gunter, LBNL
\item Alexander Konovalov, \choinote{TBD}
\item Tom Crick,  \choinote{TBD}

\end{itemize}

\subsection{Summary of Discussion}

Core questions that will need to be explored are in knowledge management, 
(transitions between people), reliability (reproducibility), usability, and how a software tool becomes part of the core workflow of well identified users (stakeholders)
relating to tool success and hence sustainability. \katznote{prev sentence is complex and awkward} Ideas 
that may need to be explored include:
\begin{itemize}
\item Requirements engineering to create tools with immediate uptake;
\item When should software ``die''?
\item Catering to disruptive developments in environment, e.g., new hardware, new methodology;
\item Dimensions of sustainability -- economic, technical, environmental, 
declining interest in primary application area), \katznote{not sure what the prev. comment goes with} social.
\end{itemize}

Sustainability requires community participation in code development and/or a wide adoption of software.
The larger the community base is using a piece of software, the better are the funding possibilities and thus also the sustainability options.
Additionally developer commitment to an application is essential and experience shows that software packages with an evangelist imposing strong inspiration and discipline are more likely to achieve sustainability.
While a single person can push sustainability to a certain level, open source software also needs sustained commitment from the developer community.
Such sustained commitments include diverse tasks and roles, which can be fulfilled by diverse developers with different knowledge levels.
Besides developing software and appropriate software management with measures for extensibility and scalability of the software, active (expertise) support for users via a user forum with a quick turnaround is crucial.
The barrier to entry for the community as users as well as developers has to be as low as possible.

\subsection{Description of Opportunity, Challenges, and Obstacles}

The opportunity lies in collaboration on a white paper, which will be revisited regularly for further improvements, to enhance knowledge of the state of best practices, resulting in a peer-reviewed paper. We would like to reach a wide community by doing this. But these are also the challenges and obstacles -- to get everyone to contribute to the paper and to reach the community.

\subsection{Key Next Steps}

The key next steps are to write the introduction, reach out to the co-authors, and to agree on the scope of the white paper.

\subsection{Plan for Future Organization}

Sandra Gesing and Abani Patra are the main editors and will organize the overall communication and paper. Sections will be assigned to diverse co-authors.

\subsection{What Else is Needed?}

At the moment we don't see any further requirements.

\subsection{Key Milestones and Responsible Parties}
\begin{itemize}
\item 15 Nov: Introduction and scope finished (Abani Patra/Sandra Gesing)
\item 15 Nov: Sections assigned (Abani Patra/Sandra Gesing)
\item 31 Jan: Analyzing funding possibilities for survey
\item 31 Jan: First versions of section
\item 15 Feb: Distribution to WSSSPE community
\item 31 Mar: Final version of white paper
\item 30 Apr: Submission of peer-reviewed paper?
\end{itemize}

\subsection{Description of Funding Needed}
We might need funding for a journal publication (open-access options).