### 2016-01-12 WSSSPE3/report/wg_main_industry_interaction.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Transition Pathways to Sustainable Software: Industry \& Academic Collaboration}
\label{sec:industry_interaction}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\subsubsection{Why it is important}

Most scientific software is produced as a part of grant-funded research projects
sponsored by federal governments. If we are interested in the sustainability of
scientific software, then we need to understand what exactly happens when that
sponsorship ends. More than likely, the project and its resulting software will
need to undergo some kind of transition in funding and consequently management.

At WSSSPE3 our group was interested in better understanding successful pathways
for scientific software to ``transition'' from government-funded research
projects to industry sponsorship. (This may be an initially awkward
phrase---some software projects will begin their life being sponsored by
industry, or result in collaboration between industry and academia. In such
cases, there is still a need to understand how IP and how maintenance of the
software is sustained over time.)

\subsubsection{Fit with related activities}

Most previous research and discussion of industry and academic collaboration, sharing, and funding of research software has focused on the impact of such arrangements. Examples of these types of reports are as follows:

\begin{itemize}
\item REF Impact Case Studies: \url{http://impact.ref.ac.uk/CaseStudies/}
\item Background of projects funded in the UK: \url{http://gtr.rcuk.ac.uk/}
\item Dowling Review from the UK: addresses complexity of work between these two
communities: \url{http://www.raeng.org.uk/policy/dowling-review}
\item Pathway to Impact - UK report: two pages of grant proposals are asked to
forecast what impact they might have (including environmental, academic, economic).
\end{itemize}

\subsubsection{Discussion}

Although sustainability transitions are often studied under the broad umbrella of ``technology
transfer,'' we believe there are likely to be a number of different ways in which
a pathway from initial production to long-term maintenance and secure funding is
achieved. In short, industry sponsorship is an important aspect of sustaining
scientific software, but our current understanding of these transitions focuses
narrowly on commercial successes or failures of those collaborations.

In looking at existing literature that addresses industry transitions, many
reports (such as those listed above) focus on benefits that accrue to the private sector, or to a government
that originally sponsored the research project. This literature does not address
the impact that these transitions have on the accessibility or usability of the
software, or the impact that these transitions have on the career of the
researchers involved.

For more detail on the group's discussion, see
Appendix~\ref{sec:appendix_industry_interaction}.

\subsubsection{Plans}

Plans for carrying forward are currently unclear---this project would require
sustained attention and effort from our team, and at least some amount of
funding in order for us to be involved for extended periods of time.

The broad goals that we would like to accomplish are:

\begin{enumerate}
\item To complete a set of case studies which look at successful and unsuccessful transitions between academic researchers and industry
\item To create a generalizable framework, which might allow for a broader study of different transition pathways (other than between academia and industry)
\end{enumerate}

The main plan for the group going forward is the creation of a white paper on the topic of sustainability transitions.

\subsubsection{Landing Page}

In lieu of a landing page, the Transitions Pathways
group requests an email be sent to Nic Weber\footnote{email: 
\href{mailto:nmweber@uw.edu}{nmweber@uw.edu}} to find out more
about the group's efforts and how to participate.