### 2016-02-03 WSSSPE3/report/wg_main_best_practices.tex
Section 5.1: Add line breaks. Add \medskip for better spacing.

\subsection{White paper/journal paper about best practices in developing sustainable software}
\label{sec:best-practices}

%\subsubsection{Why it is important}
Reviewing multiple past articles and talks at different meetings like WSSSPEx
\cite{WSSSPE1, WSSSPE2, 5069157, Blatt_WSSSPE, Ahern_WSSSPE} and analyzing and
promoting sustainable scientific software makes it clear that there are several
common and recurring ideas that underpin success in developing sustainable
software. However, outside of a small community, this knowledge is not widely
shared. This is especially true for the large community of scientists who
generate most of the software used by scientists but are not primarily software
developers. In this scenario, a clear and precise exposition of these best
practices collected from many sources and open collaboration among all in the
community in a single source (e.g., journal paper, tutorial) that can be widely
disseminated is necessary and likely to be very valuable.

\subsubsection{Fit with related activities}

The creation of such a ``best practices'' document will build upon the range of
activities and topics discussed at WSSSPE3 and associated prior meetings. We
will attempt to distill the emerging body of knowledge into this document. The
large number of articles from the NSF funded SI2 projects (SSE and SSI),
``lightning talks'', ``white papers,'' and reports from different workshops have
created a large if somewhat diffuse source for this report.

\subsubsection{Discussion}

Core questions that will need to be explored are in reliability
(reproducibility), usability, extensibility, knowledge management and continuity
(transitions between people). Answers to these will guide us on how a software
tool becomes part of the core workflow of well identified users (stakeholders)
relating to tool success and hence sustainability.
%\katznote{prev sentence is complex and awkward} % see fix
Ideas  that may need to be explored include:
\begin{itemize}
\item Requirements engineering to create tools with immediate uptake;
\item When should software ``die''?
\item Catering to disruptive developments in environment (e.g., new hardware,
new methodology);
\item Dimensions of sustainability: economic, technical, environmental and
obsolescence. 
%\katznote{not sure what the prev. comment goes with} %social.
\end{itemize}

Sustainability requires community participation in code development and/or a
wide adoption of software. The larger the community base is using a piece of
software, the better are the funding possibilities and thus also the
sustainability options. Additionally developer commitment to an application is
essential and experience shows that software packages with an evangelist
imposing strong inspiration and discipline are more likely to achieve
sustainability. While a single person can push sustainability to a certain
level, open source software also needs sustained commitment from the developer
community. Such sustained commitments include diverse tasks and roles, which can
be fulfilled by diverse developers with different knowledge levels. Besides
developing software and appropriate software management with measures for
extensibility and scalability of the software, active (expertise) support for
users via a user forum with a quick turnaround is crucial. The barrier to entry
for the community as users as well as developers has to be as low as possible.

For additional information about the discussion, see
Appendix~\ref{sec:appendix_best_practices}.

\subsubsection{Plans}
%\todo{short text here - not bullets}
The creation of a best-practices document needs a large and diverse community
involved. We have enlisted over ten contributors from the attendees at the
WSSSPE3 and those on the mailing list. The primary mechanism for developing this
document will be to examine and analyze the success of several well known
community scientific software and organizations supporting scientific software.
We will attempt then to abstract general principles and best practices. Some of
the tools identified for such analysis are the general purpose PETSc toolkit for
linear system solution, NWChem for computational chemistry and the CIG
(Computational Infrastructure for Geodynamics) organization dedicated to
supporting an ensemble of related tools for the geodynamics community. We also
established a timeline and a rough outline (see
Appendix~\ref{sec:appendix_best_practices}) for the report.

\medskip
\noindent{\bf Timeline:}
\begin{itemize}
%
\item 28 Dec: Introduction and scope finished
\item 06 Jan: Sections assigned
\item 31 Jan: Analyzing funding possibilities for survey
\item 31 Jan: First versions of section
\item 15 Feb: Distribution to WSSSPE community
\item 31 Mar: Final version of white paper
\item 30 Apr: Submission of peer-reviewed paper?
\end{itemize}

\subsubsection{Landing Page}
The landing page with instructions, timeline and the white paper is here: \url{https://drive.google.com/drive/folders/0B7KZv1TRi06fbnFkZjQ0ZEJKckk}.
Discussions can be also continued in \url{https://github.com/WSSSPE/meetings/issues/42}.