### 2016-02-05 WSSSPE3/report/wg_main_eng_design.tex
Section 5.5 and Appendix G: spacing.

\subsection{Principles for Software Engineering Design for Sustainable Software}

%\subsubsection{Why it is important}

Software engineering principles form the basis of methods, techniques,
methodologies and tools~\cite{Vliet:2008}. However, there is often a mismatch
between software engineering theory and practice particularly in the fields of
computational science and engineering, which can lead to the development of
unsustainable software~\cite{Merali:2010,hettrick:2014}. Understanding and
applying software engineering principles is essential in order to create and
maintain sustainable software~\cite{Becker:2016}.

\subsubsection{Fit with related activities}

The group discussion focused on identifying existing principles of software
engineering design that could be adopted by the computational science and
engineering communities.

\subsubsection{Discussion}

Software engineering principles form the basis of methods, techniques,
methodologies and tools. This group, which included members from different
backgrounds, including quantum chemistry, epidemiology, computer science,
software engineering, and microscopy, discussed the principles of software
engineering design for sustainable software (starting with principles from the
Karlskrona Manifesto on Sustainability Design~\cite{Becker:2015},
Tate~\cite{tate:2005}, and the Software Engineering Body of Knowledge
(SWEBOK)~\cite{swebokv3}) and their application in various domains including
quantum chemistry and epidemiology. The group examined the principles and took a
retrospective analysis of what the developers did in practice against how the
principles could have made a difference, and asked, what do the principles mean
for computational scientific and engineering software, and how do the principles
relate to non-functional requirements? It appeared that the sustainable software
engineering principles should be mapped to two core quality attributes that
underpin technically sustainable software: extensibility, the software's ability
to be extended and the level of effort required to implement the extension; and
maintainability: the effort required to locate and fix an error in operational
software.

For more information about the discussion, see Appendix~\ref{sec:appendix_eng_design}.

\subsubsection{Plans}

The next steps in this endeavor are to (1) Systematically analyze a number of
example systems from different scientific domains with regards to the identified
principles, to (2) Identify the commonalities and gaps in applying those
principles to different scientific systems, and to (3) Propose a set of
guidelines on the principles and how they exemplary apply to scientific software
system. Preliminary work will be carried out through undergraduate or
post-graduate student projects.

\subsubsection{Landing Page}

In the absence of a landing page, the Principles for Software Engineering Design
for Sustainable Software working group requests an email be sent to Birgit
Penzenstadler\footnote{email:
\href{mailto:birgit.penzenstadler@csulb.edu}{birgit.penzenstadler@csulb.edu}}
and Colin C.\
Venters\footnote{email:\href{mailto:c.venters@hud.ac.uk}{c.venters@hud.ac.uk}}
to find out more about the group's efforts and how to participate.