### 2016-02-05 WSSSPE3/report/wg_appendix_eng_design.tex
Section 5.5 and Appendix G: tiny changes.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Engineering Design Group Discussion}
\label{sec:appendix_eng_design}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Birgit Penzenstadler\footnote{email:
\href{mailto:birgit.penzenstadler@csulb.edu}{birgit.penzenstadler@csulb.edu}}
and Colin C.\ Venters\footnote{email:
\href{mailto:c.venters@hud.ac.uk}{c.venters@hud.ac.uk}} will serve as the points
of contact for this working group, and be responsible for ensuring timely
progress of the planned actions.

\subsection{Group Members}

\begin{itemize}
\item Birgit Penzenstadler -- California State University, CA, USA
\item Colin C.\ Venters -- University of Huddersfield, Huddersfield, UK
\item Matthias Bussonnier -- UC Berkeley, CA, USA
\item Jeff McWhirter -- Geode Systems 
\item Patrick Nichols -- National Center for Atmospheric Research, CO, USA
\item Ilian Todorov -- Science \& Technology Facilities Council, UK
\item Ian Taylor -- Cardiff University, UK
\item Alexander Vyushkov -- University of Notre Dame, IN, USA
\end{itemize}

\subsection{Summary of Discussion}

This group is comprised  of members from different backgrounds, including quantum
chemistry, epidemiology, microscopy, computer science, and software engineering.
Each participant was invited to give their perspective on the topic area and
what they thought were the crucial points for discussion. There was a general
consensus that there was a need for relating principles to practice for the
computational science and engineering community. Furthermore, various members of
the group expressed their interest in tools and best practices for facilitating
the maintenance and evolution of scientific software systems. It was agreed to
identify principles from software engineering and from sustainability design
and, based on those lists, discuss what each of those would mean applied to
specific example systems from the expert domains of some of the group members.
The group identified a number of software engineering principles drawn from the
Software Engineering Body of Knowledge (SWEBOK)~\cite{swebokv3}.

Software design principles include abstraction; coupling and cohesion;
decomposition and modularization; encapsulation and information hiding;
separation of interface and implementation; sufficiency completeness and
primitiveness; and separation of concerns. Similarly, user interface design
principles include learnability; user familiarity; consistency; minimal
surprise; recoverability; user guidance; and user diversity. The sustainability
design principles were drawn from the Karlskrona Manifesto on Sustainability
Design~\cite{Becker:2014}. The manifesto states that sustainability is systemic,
multidimensional, and interdisciplinary; transcends the system's purpose; applies to
both a system and its wider contexts; requires action on multiple levels;
requires multiple timescales; changing design to take into account long-term
effects does not automatically imply sacrifices; system visibility is a
precondition for and enabler of sustainability design. A number of sustainable
software engineering principles proposed by Tate~\cite{tate:2005} were also
considered including: continual refinement of product and project practices; a
working product at all times; continual emphasis on design; and value defect
prevention over defect detection.

This congregated list is an initial collection of principles that could be
extended by adding from further related work form separate disciplines within
the field of software engineering, including requirements engineering, software
architecture, and testing. The group identified two example systems to discuss
the application of the principles. The first one was a quantum chemistry system
that allows the analysis of the characteristics and capabilities of molecules
and solids. The second one was a modeling system for malaria that permitted
biologists to analyze a range of datasets across geography, biology, and
epidemiology, and add their own datasets. The group then examined the principles
and took a retrospective analysis of what the developers did in practice against
how the principles could have made a difference. This raised the question, what
do the principles mean for computational scientific and engineering software?
Similarly, how do the principles relate to non-functional requirements? It was
suggested that at the very minimum, that sustainable software engineering
principles should be mapped to two core quality attributes that underpin
technically sustainable software:
\begin{itemize}
\item Extensibility: the software's ability to be extended and the level of
effort required to implement the extension;

\item Maintainability: the effort required to locate and fix an error in
operational software.

\end{itemize}
These fundamental building blocks could then be extended to include other
quality attributes such as portability, reusability, scalability, usability, and
energy efficiency etc. Nevertheless, this raises the question of what metrics
and measures are suitable to demonstrate the sustainability of the software. In
addition, what do the five dimensions of sustainability mean for scientific
software, i.e., environmental, economic, social, technical and individual?

\subsection{Description of Opportunity, Challenges, and Obstacles}
The opportunity was identified to distill existing software engineering and
sustainability design knowledge into ``bite sized'' chunks for the Computational
Science and Engineering Community. In addition, two primary challenges were
identified:
\begin{itemize}
\item Mapping of the principles to best practices.
\item Demonstrating the return on investment of those best practices.
\end{itemize}

\subsection{Key Next Steps}

In order to achieve the following three goals: (1) a systematic analysis of a number of example systems from
different scientific domains with regards to the identified principles, (2) the
identification of the commonalities and gaps in applying the principles to different
scientific systems, and (3) a proposal of a set of guidelines on the
principles, the next steps were discussed in the following subsections.

\subsection{Plan for Future Organization}

The following plan for future organization was discussed:
\begin{itemize}
\item Identify suitable undergraduate or post-graduate students.
\item Design and pilot study.
\item Organize coordinating online calls via Google Hangout.
\end{itemize}

\subsection{What Else is Needed?}

\begin{itemize}
\item Ethics committee review panel approval required for data collection.
\end{itemize}

\subsection{Key Milestones and Responsible Parties}

The following key milestones were discussed as a roadmap for the set of
guidelines on software engineering principles:
\begin{itemize}
\item Oct/Nov 2015: Study design and interview guideline
\item Jan/Feb 2016: Interviews conducted and transcribed
\item Mar/Apr 2016: Analysis complete
\item May 2016: Report written
\end{itemize}

\subsection{Description of Funding Needed}

Specific funding was not discussed in this working group. However, this is a
open topic that can be discussed in relation to emerging funding calls from
National agencies or grant proposal initiatives.