### 2016-08-31 WSSSPE3/report/wg_main_funding_spec_expert.tex
changing subsections of 5 to not be first person

\subsection{Funding Research Programmer Expertise}
\label{RSE}

%\katznote{I don't think this section title is exactly right - are there some other options?  Also, the title of Appendix~\ref{sec:appendix_funding_spec_expert} should match, whatever is chosen.  Maybe `Funding Research Programmer Expertise'?}

%\subsubsection{Why it is important}

Research Software Engineers (RSEs)---those who contribute to science and
scholarship through software development---are an important part of the team
needed to deliver 21st century research. However, existing academic structures
and systems of funding do not effectively fund and sustain these skills. The
resulting high levels of turnover and inappropriate incentives are significant
contributing factors to low levels of reliability and readability observed in
scientific software. Moreover, the absence of skilled and experienced developers
retards progress in key projects, and at times causes important projects to fail
completely.

Effective development of software for advanced research requires that
researchers work closely with scientific software developers who understand the
research domain sufficiently to build meaningful software at a reasonable pace.
This requires a collaborative approach---where developers who are fully engaged
or invested in the research context are co-developing software with domain
academics.

\subsubsection{Fit with related activities}

The solution this group envisions entails creating an environment where software
developers are a stable part of a research team. Such an environment mitigates
the risk of losing a key developer at a critical moment in a projects lifetime,
and provides the benefits of building a store of institutional knowledge about
specific projects as well as about software development for today's research.
The group's vision is to find a way to promote a university/research institute
environment where software developers are stable components of research project
teams.

One strategy to promote stability is implementing a mechanism for developers to
obtain academic credit for software development work (see \S\ref{sec:software-credit}.) With such a mechanism in
place, traditional academic funding models and career tracks could properly
sustain individuals for whom software development is their primary contribution
to research. A contributing factor to the problem with the current academic
reward system is the devastating effect on an academic publication record
resulting from time in industry; such postings often develop exactly the skills
that research software engineers need, yet returns to university positions
following an industry role are penalized by the current structures. Retention of
senior developers is hard, because these people are high in demand by the
economy. However, people who have a PhD in science and enter industry, may
desire to return for diverse reasons, and should be welcomed back.

While developing new mechanisms in the current academic reward system is a
worthy aspirational goal, such a dramatic change in this structure does not seem
likely in a time scale relevant to this working group. Accordingly, the working
group sought alternative solutions that may be achievable within the context of
existing academic structures. The group felt that developing dedicated research
software engineering roles within the university and finding stable funding for
those individuals is the most promising mechanism for creating a stable software
development staff.

Measures of impact and success for research programming groups, as well as for
individual research software engineers, will be required in order to make the
case to the university for continued funding. Research software engineers will
hopefully not be measured by publications, but by other metrics. Middle-author
publications are common for RSEs. Most RSEs welcome co-authorship on papers
when the PIs think that the contribution deserves it.

\subsubsection{Discussion}

It is hard for an individual PI in a university or college to support dedicated
research software engineering resources, as the need and funding for these
activities are intermittent within a research cycle. To sustain this capacity,
therefore, it is necessary to aggregate this work across multiple research
groups.

One solution is to fund dedicated software engineering roles for major research
software projects at national laboratories or other non-educational
institutions. This solution is in place and working well for many well-used
scientific codebases. However, this strategy has limited application, as much of
the body of software is created and maintained in research universities.
Therefore, the group argues that research institutions should develop hybrid
academic-technical tracks for this capacity, where employees in this track work
with more than one PI, rather than the traditional RA role within a single
group. This could be coordinated centrally, as a core facility, perhaps within
research computing organizations which have traditionally supported university
cyberinfrastructure, library organizations, or research offices. Alternatively,
these groups could be organizationally closer to research groups, sitting within
academic departments. The most effective model will vary from institution to
institution, but the mandate and ways of working should be similar.

Having convinced themselves that this would be a positive innovation, the group members were
then faced with the specific question of how to fund the initiation of this
activity. A self-sustaining research software group will support itself through
collaborations with PIs in the normal grant process, with PIs choosing to fund
some amount of research software engineering effort through grants in the usual
way. However, to bootstrap such a function to a level where it has sufficient
reputation and client base to be self-sustaining will generally require seed
investment.

This might come from universities themselves (this was the model that led to the
creation of the group in University College London), but more likely, seed
funding needs to come from research councils or other funding bodies (as with the Research Software
Engineering Fellowship provided by the UK Engineering and Physical Sciences
Research Council). The group therefore recommends that funding organizations consider
how they might provide such seed funding.

Success, appropriately measured, will help make the case to such funding bodies
for further investment. One might expect that metrics such as improved
productivity, software adoption rates, and grant success rates would be
sufficient arguments in favor of such a model. However, useful measurement of
code cleanliness, and the resulting productivity gains, is an unsolved problem
in empirical software engineering. To measure ``what did not go wrong'' because
of an intervention is particularly hard.

The working group finally noted that the institutional case for such groups is made easier by
having successful examples to point to. In the UK, a collective effort to
identify the research software engineering community, with individuals clearly
stating ``I am a research software engineer,'' has been important to the
campaign. It will be useful to the global effort to similarly identify emerging
research software organizations, and also, importantly, to identify
longer-running research software groups, which have in some cases had a long
running \emph{sui-generis} existence, but which now can be identified as part of
a wider solution. There remains the problem of how to ``sell'' the value of this
investment to investigators within a university. This is an issue best addressed
by the individual organizations that embark on the plan.

For more details on the discussion, see
Appendix~\ref{sec:appendix_funding_spec_expert}.

\subsubsection{Plans}

The first step in moving this strategy forward is to gather a list of groups
that self-identify as research software engineering groups, and to reach out to
other organizations to see if there may be a widespread community of RSEs who do
not identify themselves as such at this time. This working group will collect information about
the organizational models under which these groups function, and how they are
funded. For example, how many research universities currently fund people in the
RSE track, whether they bear the RSE moniker or not. Are these developers paid
by the university or through a program supported by research grants/individual
PIs? How did they bootstrap the developer track to get this started? How
successful is the university in getting investigators to pay for fractional
RSEs? The group will author a report describing their findings, should funding be
available to conduct the investigation.

\subsubsection{Landing Page}

A list of known UK research software engineering groups is available at  
\url{http://www.rse.ac.uk/groups}, and a list for the rest of the world
is at \url{http://www.rse.ac.uk/international}.
To add another group to the list, please make a pull request as requested
on either of these pages.