### 2016-09-14 WSSSPE4_report/orgcom.tex
added organizing committee

Gabrielle Allen &  University of Illinois Urbana--Champaign, USA \\
Jeffrey Carver  & University of Alabama, USA \\
Sou-Cheng (Terrya) Choi &  Illinois Institute of Technology, USA \\
Tom Crick & Cardiff Metropolitan University, UK \\
Michael R.~Crusoe & Common Workflow Language Project \\
Sandra Gesing & University of Notre Dame, USA \\
Robert Haines & University of Manchester, UK \\
Michael Heroux & Sandia National Laboratories, USA \\
Lorraine J.~Hwang & University of California, Davis, USA \\
Daniel S.~Katz & University of Illinois Urbana--Champaign, USA \\
Kyle E.~Niemeyer & Oregon State University, USA \\
Manish Parashar & Rutgers University, USA \\
Colin C.~Venters & University of Huddersfield, UK