### 2016-09-14 WSSSPE4_report/awardees.tex
added superscripts for students and early-career awardees

Alice Allen & Astrophysics Source Code Library \\
Wolfgang Bangerth & Colorado State University, USA \\
Roman Bauer\textsuperscript{\textdagger} & Newcastle University, UK \\
Steven Brandt & Louisiana State University, USA \\
Emily Chen* & University of Illinois Urbana-Champaign, USA \\
Bruce Childers & University of Pittsburgh, USA \\
Michael R.~Crusoe\textsuperscript{\textdagger} & Common Workflow Language Project \\
Stephan Druskat\textsuperscript{\textdagger} & Humboldt-Universit\"{a}t zu Berlin, Germany \\
Anshu Dubey & Argonne National Laboratory, USA \\
Iain Emsley* & Oxford University, UK \\
Christopher Gwilliams\textsuperscript{\textdagger} & Cardiff University, UK \\
Lorraine Hwang & University of California Davis, USA \\
Ray Idaszak & RENCI, University of North Carolina at Chapel Hill, USA \\
Xinlian Liu & Hood College, USA \\
Frank Löffler & Louisiana State University, USA \\
Tengyu Ma* & Vanderbilt University, USA \\
Robert H.~McDonald & Indiana University, USA \\
Jonah Miller* & Perimeter Institute, Canada \\
Daniel Mosse & University of Pittsburgh, USA \\
Kyle E.~Niemeyer\textsuperscript{\textdagger} & Oregon State University, USA \\
Santiago N\'{u}\~{n}ez-Corrales* & University of Illinois Urbana-Champaign, USA \\
Dominic Orchard & University of Kent, UK \\
Francisco Queiroz* & Pontifical Catholic University of Rio de Janeiro, Brazil \\
Jean Salac* & University of Virginia, USA \\
Janos Sallai & Vanderbilt University, USA \\
Prabhjyot (Singh) Saluja* & Southwest Oklahoma State University, USA \\
Eric Seidel* & University of California San Diego, USA \\
Karan Shah* & Georgia Institute of Technology, USA \\
Sameer Shende* & University of Oregon, USA