### 2016-09-14 WSSSPE4_report/awardees.tex
changed * to superscript in awardees list

Alice Allen & Astrophysics Source Code Library \\
Wolfgang Bangerth & Colorado State University, USA \\
Roman Bauer\textsuperscript{\textdagger} & Newcastle University, UK \\
Steven Brandt & Louisiana State University, USA \\
Emily Chen\textsuperscript{*} & University of Illinois Urbana-Champaign, USA \\
Bruce Childers & University of Pittsburgh, USA \\
Michael R.~Crusoe\textsuperscript{\textdagger} & Common Workflow Language Project \\
Stephan Druskat\textsuperscript{\textdagger} & Humboldt-Universit\"{a}t zu Berlin, Germany \\
Anshu Dubey & Argonne National Laboratory, USA \\
Iain Emsley\textsuperscript{*} & Oxford University, UK \\
Christopher Gwilliams\textsuperscript{\textdagger} & Cardiff University, UK \\
Lorraine Hwang & University of California Davis, USA \\
Ray Idaszak & RENCI, University of North Carolina at Chapel Hill, USA \\
Xinlian Liu & Hood College, USA \\
Frank Löffler & Louisiana State University, USA \\
Tengyu Ma\textsuperscript{*} & Vanderbilt University, USA \\
Robert H.~McDonald & Indiana University, USA \\
Jonah Miller\textsuperscript{*} & Perimeter Institute, Canada \\
Daniel Mosse & University of Pittsburgh, USA \\
Kyle E.~Niemeyer\textsuperscript{\textdagger} & Oregon State University, USA \\
Santiago N\'{u}\~{n}ez-Corrales\textsuperscript{*} & University of Illinois Urbana-Champaign, USA \\
Dominic Orchard & University of Kent, UK \\
Francisco Queiroz\textsuperscript{*} & Pontifical Catholic University of Rio de Janeiro, Brazil \\
Jean Salac\textsuperscript{*} & University of Virginia, USA \\
Janos Sallai & Vanderbilt University, USA \\
Prabhjyot (Singh) Saluja\textsuperscript{*} & Southwest Oklahoma State University, USA \\
Eric Seidel\textsuperscript{*} & University of California San Diego, USA \\
Karan Shah\textsuperscript{*} & Georgia Institute of Technology, USA \\
Sameer Shende\textsuperscript{*} & University of Oregon, USA