### 2016-10-05 WSSSPE4_report/wg_open_research_index.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Open Research Index}
\label{sec:open-research-index}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\note{Dan to write this}

Introduction to group here, including the overall objective of work in this area.

\subsubsection{Participants}

members of the working group

\subsubsection{Working group objective}

Specific things the working group wants to accomplish in the context of the larger objective.

\subsubsection{Gap or challenge}

What is the gap or challenge being addressed?

\subsubsection{Relevant people and resources}

What people, groups, or resources are needed.

\subsubsection{Plans}

What tasks will the working group undertake

\subsubsection{SMART steps}

What are the first SMART steps proposed?

\subsubsection{More information \& joining instructions}

How could a reader get more information or get more involved?