### 2016-11-07 WSSSPE4_report/wg_codemeta.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{CodeMeta}
\label{sec:CodeMeta}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\note{Alice to write this}

Introduction to group here, including the overall objective of work in this area.

Something in here about this group having formed in reaction to one of the project suggestions made on the second day of the meeting.

On the second day of WSSSPE4, participants were challenged to share their ideas for ideal future world, list gaps that exist between now and future possibilities, and generate possible projects to bridge existing gaps to move us closer to the ideal vision of attendees. One of the possible projects listed was to "Define a standard for metadata to be used in current repos including authors, dependencies, ... repo url." Several attendees have been participating in a project, CodeMeta, that seeks to do just that. Interest in this possible was very high and an enthusiastic group formed with the object of fostering adoption of CodeMeta.

\subsubsection{Participants}

Gabrielle Allen <gdallen@illinois.edu>
Stephan Druskat <stephan.druskat@hu-berlin.de>
Iain Emsley <iain.emsley@oerc.ox.ac.uk>
Carole Goble <carole.goble@manchester.ac.uk>
Chris Gwilliams <gwilliamsc@cardiff.ac.uk>
Rafael Jimenez <rafael.jimenez@elixir-europe.org>
Frank Löffler <knarf@cct.lsu.edu>
Kyle Niemeyer <Kyle.Niemeyer@oregonstate.edu>
Thomas Robitaille <thomas.robitaille@gmail.com>
Rob Welch <py12rw@leeds.ac.uk>
Alice Allen <aallen@ascl.net>

\subsubsection{Working group objective}

Engage existing members with WSSSPE4 CodeMeta Working Group
Crosswalk Schema.org types 
Determine how to engage the community and provide ongoing social engagement and structure
Assist in implementation of the specifications and brush up documentation 
Define a list of services to be added and match terms (add to csv) 
Provide an outsider’s view and contribute suggestions for more understandable docs 
Find better ways to present crosswalk table

\subsubsection{Gap or challenge}

What is the gap or challenge being addressed?

\subsubsection{Relevant people and resources}

What people, groups, or resources are needed.

Original CodeMeta members
Create mailing list for WSSSPE CodeMeta participants
Engage through email

Useful links
\url{https://github.com/codemeta/codemeta}
\url{https://github.com/codemeta/codemeta/milestones}
\url{https://gitter.im/codemeta/codemeta}

\subsubsection{Plans}

What tasks will the working group undertake

\subsubsection{SMART steps}

Write to managers of CodeMeta project to start dialog with the two groups - Alice (done)
Post questions generated by our discussion as issues on CodeMeta repo - Iain, Rafa, Stephan, others as needed (done)
Crosswalk Schema.org types - Rafa, Carole
Assist in implementation of the specs and brush up documentation - Rob
Define a list of services to be added and match terms (add to csv) - Chris
Identify two roles and information on CodeMeta that would be useful to these people for engaging with the project - Stephan (done)
Think about better ways to present crosswalk table - Tom, Frank
Create mailing list for WSSSPE CodeMeta participants - Kyle (30 Sep 16)

\subsubsection{More information \& joining instructions}

Because of the work done at WSSSPE4, the CodeMeta project README file\citep{\url{https://github.com/codemeta/codemeta/blob/master/README.md}} was greatly expanded to include a description of the project that is geared to those with little or no prior knowledge of the project and including a list of contributors, information on how one can get involved, a bit of the project history and who is managing the project, and links to additional information.