### 2017-01-14 WSSSPE4_report/wg_social.tex
adding social science working group draft

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Social science for scientific software}
\label{sec:social}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\note{Stuart to write this}

Introduction to group here, including the overall objective of work in this area. \todo{}

\subsubsection{Participants}

Stuart Geiger <stuart@stuartgeiger.com>, Lorraine Hwang <ljhwang@ucdavis.edu>, Robert McDonald <rhmcdona@indiana.edu>

\subsubsection{Working group objective}

To give social scientists and practitioners working on scientific software, software sustainability, and open source and a better understanding of each other's work, as well as help them connect and coordinate on specific projects of mutual interest.

\subsubsection{Gap or challenge}

There is more and more academic research being done on topics related to software sustainability, including work on software engineering practices and management of open source projects. However, academic research in general is often siloed for many reasons, and work on topics relevant to software sustainability is no exception. There are many academic studies and projects which may be relevant for practitioners working in this area, and many projects and initiatives by practitioners that may be relevant for social scientists. However, there is a gap between these two research and practice. Furthermore, we also recognize that academic social science and research software engineering are not monolithic, and there is a need to connect people who care about research-driven best practices inside of these two domains with each other as well.

\subsubsection{Relevant people and resources}

People: a network of invested social scientists and research software engineers who care about research on software engineering practices
Resources: a mailing list and a central repository for collecting literature and documenting lessons learned

\subsubsection{Plans}

What tasks will the working group undertake \todo{}

\subsubsection{SMART steps}

\begin{itemize}
\item Collect research questions RSEs have for social scientists (done)
\item Create a mailing list for people interested in this topic (done) \url{https://groups.google.com/forum/#!forum/researchsoftwarestudies}
\item Create a central repository for collecting literature and documenting lessons learned
\item Publicize mailing list and collect (more) research questions/topics/literature/researchers
\item Write up synthesis document 
\end{itemize}

\subsubsection{More information \& joining instructions}

Join the mailing list at \url{https://groups.google.com/forum/#!forum/researchsoftwarestudies}