### 2017-04-11 WSSSPE4_report/program_committee.tex
more name fix

David Abramson & University of Queensland, Australia\\
Lorena A. Barba & George Washington University, USA\\
Ross Bartlett & Sandia National Laboratories, USA\\
Christoph Becker & University of Toronto, Canada\\
Ewout van den Berg & IBM Watson, USA\\
David Bernholdt & Oak Ridge National Laboratory, USA\\
Stefanie Betz & Karlsruhe Institute of Technology, Germany\\
Coral Calero & Universidad Castilla La Mancha, Spain\\
Ishwar Chandramouli & National Cancer Institute, National Institutes of Health, USA\\
Ruzanna Chitchyan & University of Leicester, UK\\
Karen Cranston & Duke University, USA\\
Ewa Deelman & Information Sciences Institute, University of Southern California, USA\\
Charlie E. Dibsdale & O-Sys, Rolls Royce PLC, UK\\
Anshu Dubey & Argonne National Laboratory, USA\\
Nadia Eghbal & Independent Researcher (via Ford Foundation), USA\\
Peter Elmer & CERN, Switzerland\\
Martin Fenner & DataCite, Germany\\
David Gavaghan & University of Oxford, UK\\
Mike Glass & Sandia National Laboratories, USA\\
Carole Goble & University of Manchester, UK\\
Joshua Greenberg & Alfred P. Sloan Foundation, USA\\
Michael K Griffiths & University of Sheffield, UK\\
Sarah Harris & University of Leeds, UK\\
James Hetherington & University College London, UK\\
Fred J. Hickernell & Illinois Institute of Technology, USA\\
Neil Chue Hong & Software Sustainability Institute, University of Edinburgh, UK\\
Caroline Jay & University of Manchester, UK\\
Rafael C. Jimenez & ELIXER, Cambridge, UK\\
Matthew B. Jones & University of California Santa Barbara, USA\\
Nick Jones & New Zealand eScience Infrastructure (NeSI), NZ\\
Sedef Akinli Kocak & Ryerson University, Canada\\
Jong-Suk Ruth Lee & National Institute of Supercomputing and Networking, KISTI, Korea\\
James Lin & Shanghai Jiao Tong University, China\\
Frank Löffler & Louisiana State University, USA\\
Gregory Madey & University of Notre Dame, USA\\
Ketan Maheshwari & University of Pittsburgh, USA\\
Steven Manos & University of Melbourne, Australia\\
Chris A. Mattmann & NASA JPL; University of Southern California, USA\\
Abigail Cabunoc Mayes & Mozilla Science Lab, USA\\
Robert H. McDonald & Indiana University, USA\\
Lois Curfman McInnes & Argonne National Laboratory, USA\\
Alberto Di Meglio & CERN, Switzerland\\
Chris Mentzel & Gordon and Betty Moore Foundation, USA\\
Peter Murray-Rust & University of Cambridge, UK\\
Christopher R. Myers & Cornell University, USA\\
Jarek Nabrzyski & University of Notre Dame, USA\\
Cameron Neylon & Curtin University, Australia\\
Aleksandra Pawlik & New Zealand eScience Infrastructure (NeSI), NZ\\
Fernando Perez & Lawrence Berkeley National Laboratory; University of California, Berkeley, USA\\
Marian Petre & The Open University, UK\\
Marlon Pierce & Indiana University, USA\\
Andreas Prlic & University of California, San Diego, USA\\
Karthik Ram & University of California, Berkeley, USA\\
Morris Riedel & Juelich Supercomputing Centre, Germany\\
Dave De Roure & Oxford e-Research Centre, University of Oxford, UK\\
Norbert Seyff & University of Zurich, Switzerland\\
Arfon Smith & GitHub Inc, USA\\
Borja Sotomayor & University of Chicago, USA\\
Edgar Spalding & University of Wisconsin, USA\\
Maria Spichkova & RMIT University, Australia\\
Victoria Stodden & University of Illinois Urbana--Champaign, USA\\
Matthew Turk & University of Illinois Urbana--Champaign, USA\\
Nancy Wilkins-Diehr & San Diego Supercomputer Center, University of California, San Diego, USA\\
James Willenbring & Sandia National Laboratories, USA\\
Scott Wilson & Cetis LLP, UK\\
Theresa Windus & Iowa State University and Ames Laboratory, USA\\