### 2017-04-17 WSSSPE4_report/wg_best_practices_developing.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{White paper on developing sustainable software}
\label{sec:best-practices-developing}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\note{Sandra to write this}

\note{Introduction to group here, including the overall objective of work in this area.}

\subsubsection{Participants}

\note{members of the working group}

\subsubsection{Working group objective}

\note{Specific things the working group wants to accomplish in the context of the larger objective.}

\subsubsection{Gap or challenge}

What is the gap or challenge being addressed?

\subsubsection{Relevant people and resources}

The list of contributors, as of November 10th 2016, includes (in alphabetical order): \\
Jeffrey C. Carver <carver@cs.ua.edu>,
Neil Chue Hong <N.ChueHong@software.ac.uk>,
Tom Crick <tcrick@cardiffmet.ac.uk>,
Miguel de~Val-Borro <valborro@princeton.edu>,
Hans Fangohr <fangohr@soton.ac.uk>
Sandra Gesing <sandra.gesing@nd.edu>,
Derek Groen <Derek.Groen@brunel.ac.uk>,
Dan Gunter <DKGunter@lbl.gov>,
Daniel S. Katz <d.katz@ieee.org>,
Alexander Konovalov <alexander.konovalov@st-andrews.ac.uk>,
Frank L\"offler <knarf@cct.lsu.edu>,
Suresh Marru <smarru@iu.edu>,
Kyle E. Niemeyer <kyle.niemeyer@oregonstate.edu>,
Abani Patra <abani@buffalo.edu>
and
Francisco Queiroz <chico@puc-rio.br>

\subsubsection{Plans}

\note{What tasks will the working group undertake}

\subsubsection{SMART steps}

First eight SMART steps proposed: \\
\begin{itemize}
\item List existing contributors and open GitHub for new ones - Sandra Gesing (done).
\item Establish preliminary timeline  - Sandra Gesing (done).
\item Organize related work - Francisco Queiroz (done).
\item Define scope of the paper - Sandra Gesing (done).
\item Suggest new sections - Francisco Queiroz (done).
\item Define leading authors and contributors for sections - Francisco Queiroz (done).
\item Finalize first version - Sandra Gesing (in progress - due 31st of January, 2016)
\item Distribute to WSSSPE community and collect feedback - Sandra Gesing (in progress - due 28th of February, 2016)
\end{itemize}

\subsubsection{More information \& joining instructions}

The GitHub repository for the White Paper can be found at \url{https://github.com/WSSSPE/WG-Best-Practices}. For more information and requests, join the \texttt{\#wg-best-practices} channel at WSSSPE's Slack team.