### 2017-05-03 WSSSPE4_report/participants.tex
turning off email of participants

Mark Abraham & SciLifeLab, KTH, Sweden\\
Aseel Aldabjan & University of Manchester\\
Alice Allen & Astrophysics Source Code Library\\
Gabrielle Allen & NCSA/University of Illinois\\
Pau Andrio & Barcelona Supercomputing Center\\
Wolfgang Bangerth & Colorado State University\\
Radovan Bast & UiT The Arctic University of Norway, Nordic e-Infrastructure Collaboration\\
Roman Bauer & Newcastle University\\
Steven Brandt & Louisiana State University\\
Lukas Johannes Breitwieser & CERN\\
Emily Chen & NCSA/University of Illinois Urbana--Champaign\\
Bruce Childers & University of Pittsburgh\\
Mike Croucher & University of Sheffield\\
Michael R. Crusoe & Common Workflow Language Project\\
Stephan Druskat & Humboldt-Universit\"{a}t zu Berlin\\
Anshu Dubey & Argonne National Laboratory\\
Tim Dunne & KnowInnovation\\
Peter Elmer & Princeton University\\
Iain Emsley & University of Oxford\\
Hans Fangohr & University of Southampton\\
Stuart Geiger & Berkeley Institute for Data Science\\
Sandra Gesing & University of Notre Dame\\
Carole Goble & University of Manchester, Software Sustainability Institute\\
Tom Goodale & Independent\\
Christopher Gwilliams & Cardiff University\\
Willem van Hage & Netherlands eScience Center\\
Robert Haines & University of Manchester\\
James Hetherington & University College London\\
Simon Hettrick & Software Sustainability Institute\\
Neil Chue Hong & Software Sustainability Institute\\
Lorraine Hwang & UC Davis, CIG\\
Ray Idaszak & RENCI, University of North Carolina at Chapel Hill\\
Caroline Jay & University of Manchester\\
Brian Jimenez & Barcelona Supercomputing Center\\
Rafael C Jimenez & ELIXIR Hub, UK\\
Daniel S Katz & University of Illinois Urbana--Champaign\\
Dominic Kempf & Heidelberg University\\
Frank L\"{o}ffler & Louisiana State University\\
Patricia Lago & Vrije Universiteit Amsterdam\\
Xinlian Liu & Hood College\\
Hannes Loeffler & STFC\\
Tengyu Ma & Vanderbilt University\\
Robert H McDonald & Indiana University\\
Chris Mentzel & Moore Foundation\\
Constantinos Michailidis & KnowInnovation\\
Jonah Miller & Perimeter Institute of Theoretical Physics\\
Daniel Mosse & University of Pittsburgh\\
August (Gus) Muench & American Astronomical Society\\
Santiago N\'u\~nez-Corrales & University of Illinois Urbana--Champaign\\
Udit Nangia & University of Illinois Urbana--Champaign\\
Aleksandra Nenadic & University of Manchester, SSI\\
Kyle Niemeyer & Oregon State University\\
Dominic Orchard & University of Kent\\
Francisco Queiroz & Pontifical Catholic University of Rio de Janeiro\\
Fons Rademakers & CERN\\
Karthik Ram & University of California, Berkeley\\
Sam Relton & University Of Manchester\\
Tom Robitaille & Freelance\\
Jean Salac & University of Virginia\\
Janos Sallai & Vanderbilt University\\
Prabhjyot (Singh) Saluja & Southwestern Oklahoma State University\\
Eric Seidel & University of California, San Diego\\
Sameer Shende & University of Oregon\\
Raniere Silva & University of Manchester\\
Shoaib Sufi & University of Manchester, Software Sustainability Institute\\
Chris Sweet & University of Notre Dame\\
Rob Welch & University of Leeds\\
Christopher Woods & University of Bristol\\